---
title: "RDA trait prediction tutorial"
author: "KE Lotterhos"
date: '2022-10-31'
output: html_document
---

```{r setup, include=FALSE}

libraries_needed <- c("vegan")

for (i in 1:length(libraries_needed)){
  library(libraries_needed[i],character.only = TRUE) #laptop
}

```

## Load the data

* A matrix of genotypes in 012 format (counts of reference allele)
  * number of rows = number of individuals
  * number of columns = number of SNPs
* A table with information about sampled individuals (each individual in a row)
* A table with information about SNPs (each SNP in a row)

This data was simulated in SLiM and is associated with the paper "The paradox of adaptive trait clines with non-clinal patterns in the underlying genes" by K.E. Lotterhos. Briefly, a non-Wright-Fisher model was simulated on a landscape with 6 environmental variables that reflect different aspects of thermal stress and precipitation in British Columbia. The simulation included 6 environmental traits, each of which adapted to a different environmental variable.

The `ind` table includes the xy location for each individual, the 6 exact trait values (note that these won't exactly equal the trait value calculated from the genotype matrix because of MAF filtering), and the 6 environmental values at their xy location.

The `muts` table includes the linkage group `LG`, the position of the mutation on the genetic map `pos_pyslim`, a unique ID `mutname`, the allele frequency based on the 1000 sampled individuals `a_freq_subset`, and whether or not it had effects on one or more phenotypes `causal`.

```{r load data}
setwd("~/Documents/GitHub/MVP-NonClinalAF/tutorial") #change code to your working directory

G <- read.table("Genotypes.txt")
dim(G)

ind <- read.table("Individuals.txt", header=TRUE)
dim(ind) #corresponds to rows in G
head(ind)

muts <-  read.table("SNPs.txt", header=TRUE)
dim(muts) #corresponds to columns in G
head(muts)
```

## RDA trait prediction funciton

This function predicts an environmental trait through the back-transformation of the RDA "site score" of an individual to a chosen environmental variable

```{r function}
rda_trait_pred <- function(rdaobj, env_row, K){
  #rdaobj is RDA object
  #envi row is the row of the environment in the biplot output
  #K is the number of RDA axes
  scores <- scores(rdaobj, choices=1:K)
  ind.sc <- scores$sites
  pred <- matrix(NA, nrow=nrow(ind.sc), ncol=K)
  for (k in 1:K){
    pred[,k] <- ind.sc[,k]*eigenvals(rdaobj)[k]*summary(rdaobj)$biplot[env_row,k]
  }
  trait_pred <- scale(rowSums(pred))
 return(trait_pred) 
}
```

## Example of an RDA-predicted environmental trait value

First, run the RDA
```{r example}
# Run the RDA
rdaout <- rda(t(G_full_subset) ~ subset_indPhen_df$env1_mat +
                subset_indPhen_df$env2_MTWetQ +
                subset_indPhen_df$env3_MTDQ + 
                subset_indPhen_df$env4_PDM +
                subset_indPhen_df$env5_PwarmQ +
                subset_indPhen_df$env6_PWM
              )
```

Next, check the biplot output and decide how many RDA axes to use 
```{r}
# Check the biplot output
rdaout$CCA$biplot

# Decide how many RDA axes to use in calculation
  a<- screeplot(rdaout)
  str(a)
  a$y # save this it's the eigenvalues
  prop_var <- round(a$y[1:6]/sum(a$y),3)
  cumsum(prop_var)
  plot(cumsum(prop_var), xlab="Number of RDA axes", ylab="Cumulative percent of variation explained", ylim=c(0,1))
```

Here is an example of a trait prediction for MAT using the first 3 RDA axes
```{r}
# Make the trait prediction for MAT (1st row in biplot output)
K = 3 # use 3 RDA axes to make the trait prediciton

MATtraitPredict <- rda_trait_pred(rdaout, 1, K)

# Since this is a simulation, we can compare the prediction to the true value
# Similarly, an empirical study could compare an empirically measured trait value to the RDA-predicted trait value to test how well landscape genomic data can predict functional traits
plot(ind$phenotype1_mat, MATtraitPredict)
cor(ind$phenotype1_mat, MATtraitPredict)
```