---
title: "1-AnalyzeSims"
author: "KE Lotterhos"
date: "3/12/2021"
output: html_document
editor_options: 
  chunk_output_type: console
---

setwd("~/Documents/GitHub/MVP-NonClinalAF/src")

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

packages_needed <- c("vcfR", "distances","ggplot2",  "fields", "stringr", "vegan", "robust", "mvtnorm", "viridis", "gridExtra", "devtools")

for (i in 1:length(packages_needed)){
  if(!(packages_needed[i] %in% installed.packages())){install.packages(packages_needed[i])}
}

#  require(ggplot2)}

for (i in 1:length(packages_needed)){
  library( packages_needed[i], character.only = TRUE)
}

if (!requireNamespace("BiocManager", quietly = TRUE)){install.packages("BiocManager")}
   library(BiocManager)

if (!requireNamespace("LEA", quietly = TRUE)){BiocManager::install("LEA")} 
library(LEA)

if (!requireNamespace("qvalue", quietly = TRUE)){
BiocManager::install("qvalue")
}
library(qvalue)
#help(package="LEA") 

if(!requireNamespace("ggExtra", quietly=TRUE)){
  devtools::install_github("daattali/ggExtra")
}
library(ggExtra)

.pardefault <- par()
```

```{r load data}
path <- "../sim_output_150_V2/"
outpath <- "../results/"

#seed <- 1231453
#seed <- 1231141
#seed <- 1231217
#seed <- 1231141
#seed <- 1231158
seed <- 1231157
seed <- 1231146
#2242330863068_popInfo

# Individual phenotype and fitness data in home pop
indPhen_df <- read.table(paste0(path,seed,"_ind.txt"), header=TRUE, 
                         colClasses = c("character", rep("numeric",8)))
head(indPhen_df)

pop_df <- read.table(paste0(path,seed,"_popInfo.txt"), header=TRUE, 
                         colClasses = c("character", rep("numeric",6))) 

# Local adaptation through time df
LA_df <- read.table(paste0(path,seed,"_LA.txt"), header=TRUE, 
                         colClasses = c("character", rep("numeric",10)),
                    na.strings = "NAN")
str(LA_df)
#head(LA_df)
tail(LA_df)

# Mutation stats for MAF > 0.01
muts_df <- read.table(paste0(path,seed,"_muts.txt"), header=TRUE, 
                         colClasses = c("character","numeric","character", rep("numeric",5)))
muts_df
vatemp <- muts_df$p*(1-muts_df$p)*muts_df$mutTempEffect^2
vasal <- muts_df$p*(1-muts_df$p)*muts_df$mutSalEffect^2
plot(round(vatemp/sum(vatemp),2), abs(muts_df$cor_temp), xlim=c(0,1), ylim=c(0,1))
vasal <- muts_df$p*(1-muts_df$p)*muts_df$mutSalEffect^2
if(sum(vasal)>0){
  plot(round(vasal/sum(vasal),2), abs(muts_df$cor_sal), xlim=c(0,1), ylim=c(0,1))
}



info <- read.table(paste0(path,seed,"_info.txt"), header=TRUE, 
                         colClasses = c("character", 
                                        rep("numeric",15),
                                        "character",
                                        rep("numeric",6)))
info

# VCF file from SliM
#vcf <- read.vcfR(paste0(path,seed,"_VCF_causal.vcf.gz"))
#head(vcf)
#head(vcf@fix, 50)
#dim(vcf@fix)
# example of how to find a specific mutation in the vcf file
#  muts_df[1,]
#  vcf@fix[grep(muts_df$mutID[1], vcf@fix[,"INFO"]),]
  
allsims <- read.table("0b-final_params.txt", header=TRUE)
thissim <- allsims[grep(seed, allsims$seed),]
(plotmain <- paste(thissim$level, seed, sep="\n"))

# VCF with neutral mutations
vcf_full <- read.vcfR(paste0(path,seed,"_plusneut_MAF01.recode2.vcf.gz"))
head(vcf_full)
head(vcf_full@fix, 50)
dim(vcf_full@fix)

#vcf_full@fix[15:16,]
#vcf_full@gt[15:16,1:10]
```


```{r plot subpops and migration}
mig <- read.table(paste0(path,seed,"_popInfo_m.txt"), header=TRUE)
head(pop_df)
head(mig)



par(mfrow=c(1,1), mar=c(4,4,3,1))
pdf(paste0(path,seed,"_pdf_1pop.pdf"), width=6, height=6)

if (var(mig$m)>0){
  mig_thick <- log10(mig$m) + 5
}else{
  mig_thick <- rep(1, length(mig$m))
}
plot(pop_df$x, pop_df$y, col = rgb(0,0,0,0), bty="l", xlab="x", ylab="y", main=plotmain, cex.main=0.5)
for (i in 1:nrow(mig)){
  start_x <- pop_df$x[which(pop_df$subpopID==mig$from[i])]
  start_y <- pop_df$y[which(pop_df$subpopID==mig$from[i])]
  end_x <- pop_df$x[which(pop_df$subpopID==mig$to[i])]
  end_y <- pop_df$y[which(pop_df$subpopID==mig$to[i])]
  adj = 0.01  
  if (end_x < start_x){end_x <- end_x +0.2; start_x <- start_x - 0.5}
  if (end_x > start_x){end_x <- end_x -0.2; start_x <- start_x + 0.5}
  if (end_y < start_y){end_y <- end_y +0.2; start_y <- start_y - 0.5}
  if (end_y > start_y){end_y <- end_y -0.2; start_y <- start_y + 0.5}
  arrows(start_x,start_y,end_x, end_y, col="cornflowerblue", lwd=mig_thick[i], length=0.1)
}
text(pop_df$x, pop_df$y, pop_df$N, cex=0.6)
```

```{r optimums plot}
ggplot(pop_df) + geom_point(aes(x=x, y=y,size=opt0)) + geom_point(aes(x=x, y=y, color=opt1)) + theme_classic() + scale_colour_gradient2(high=rgb(1,0.4,0.2), low="cornflowerblue", mid=rgb(0.8,0.8,0.7), name="Temp") + geom_text(aes(x=x, y=y+0.3,label=subpopID)) + labs(size="Env2") + ggtitle(plotmain) + theme(plot.title = element_text(size = 7))

plot(LA_df$gen, LA_df$cor_temp_ind, type="l", col="orange", lwd=3, main=plotmain, ylab="cor (Env, phenotype)", xlab="Generation", ylim=c(0,1))

if (sum(LA_df$cor_sal_ind, na.rm=TRUE)>0){
  points(LA_df$gen, LA_df$cor_sal_ind, col="cornflowerblue", type="l", lwd=3, lty=2)
}
legend(0,1, c("Temp", "Env2"), lwd=3, col=c("orange", "cornflowerblue"), lty=c(1,2), bty="n")  
```

First, I want to figure out which individuals are migrants. Sampling in the simulation occurred after migration but before selection, so for analysis I want to make sure to only include individuals that would have survived to reproduce.

My goal is to start with a dataset of 50 pops * 20 individuals = 1000 individuals, with approximately equal number of individuals per population.


(Part 1) I tried to subset the data to individuals who have highest fitness in their home population, but this does not give equal numbers.

(Part 2) Decision: sample X individuals from each population with probability equal to their fitness. This will make the sampling less artificial.

Note that this sampling doesn't really make a huge difference in terms of the correlation between salinity and temperature.
```{r find high fitness individuals}

### Part 1 ####
  # sum(indPhen_df[,"fitness"]>0.8)
  # sum(indPhen_df[,"fitness"]>0.9)
  # sum(indPhen_df[,"fitness"]>0.95)
  # 
  # subset <- which(indPhen_df[,"fitness"]>0.95)
  # length(subset)
  # table(indPhen_df[subset,"subpop"])
  
### Part 2 ####
  npops <- length(levels(factor(indPhen_df$subpop)))
  n = 10 # number of individual per pop
  subset <- c()
  for (i in 1:npops){
    #set.seed(139982)
    bob <- sample(indPhen_df$indID[indPhen_df$subpop==i], size = n,
                  replace=FALSE, prob = indPhen_df$fitness[indPhen_df$subpop==i])
    subset <- c(subset, bob)
  }
  length(subset)
  head(subset)
  indPhen_df$subset <- FALSE
  indPhen_df$subset[sort((subset+1))] <- rep(TRUE, length(subset))
    # add one to subset here because indID starts at 0
  
  nind_samp <- sum(indPhen_df$subset)
  # visualize all individuals
```  

```{r}
  indPhen_df$subpopID <- indPhen_df$subpop
  indPhen_df <- merge(indPhen_df, pop_df)
  
  boxplot(indPhen_df$fitness~indPhen_df$subpop, ylab = "fitness", xlab="pop", main="all", ylim=c(0,1))
  boxplot(indPhen_df$fitness~indPhen_df$N, ylab = "fitness", xlab="N", main="all", ylim=c(0,1))
  
  boxplot(indPhen_df$fitness[indPhen_df$subset]~indPhen_df$subpop[indPhen_df$subset], ylab = "fitness", xlab="pop", main="subset", ylim=c(0,1))
  
  boxplot(indPhen_df$fitness[indPhen_df$subset]~indPhen_df$N[indPhen_df$subset], xlab="N", ylab="fitness")
  # looks good to me
  dev.off()
  
  if (var(indPhen_df$phen_sal)>0){
    all_corr_phen_sal <- cor(indPhen_df$phen_sal,indPhen_df$sal_opt)
      subsamp_corr_phen_sal <- cor(indPhen_df$phen_sal[indPhen_df$subset], indPhen_df$sal_opt[indPhen_df$subset])
  }else{
    all_corr_phen_sal <- NA
    subsamp_corr_phen_sal <- NA
  }
  
  all_corr_phen_temp <- cor(indPhen_df$phen_temp,indPhen_df$temp_opt)
  
  subsamp_corr_phen_temp <- cor(indPhen_df$phen_temp[indPhen_df$subset],indPhen_df$temp_opt[indPhen_df$subset])
```


```{r show cool patterns of mutation}

pdf(paste0(path,seed,"_pdf_2muts.pdf"), width=5, height=5)

plot(muts_df$mutSalEffect, muts_df$mutTempEffect, ylim=c(-0.7, 0.7), xlim=c(-0.7,0.7), xlab="mutation effect on Env2 phenotype", ylab="mutation effect on temperature phenotype", bty="l", main=plotmain)
abline(h=0, col="grey")
abline(v=0, col="grey")

```


```{r VCF_FULL keepmuts}
geno_full <- vcf_full@gt[,-1] 
dim(geno_full)
position_full <- getPOS(vcf_full)
rown <- as.numeric(vcf_full@fix[,"ALT"]) # mutation ID
  # rown = 1 is a neutral mutation
causal_mut_locs <- which(rown %in% muts_df$mutID) # causal mutations
position_full[causal_mut_locs] # positions of causal mutations
```


The following code is for the subset VCF file of ALL mutations, subset to sampled individuals
```{r filter VCF_FULL and reformat}

head(geno_full[,1:5])


G_full <- matrix(NA, nrow = nrow(geno_full), ncol = ncol(geno_full))
G_full[geno_full %in% c("0/0", "0|0")] <- 0
G_full[geno_full  %in% c("0/1", "1/0", "1|0", "0|1")] <- 1
G_full[geno_full %in% c("1/1", "1|1")] <- 2
rm(geno_full)


head(G_full[,1:5])

a_freq_full <- rowSums(G_full)/(2*ncol(G_full))
hist(a_freq_full)

if (!identical(sort(unique(unlist(as.numeric(G_full)))), as.numeric(0:2))){
  print("Error: full genotype matrix not uniquely 012")
  break
}

G_full_subset <- G_full[,indPhen_df$subset]
rm(G_full)

dim(G_full_subset)

a_freq_subset <- as.numeric(rowSums(G_full_subset)/(2*ncol(G_full_subset)))
  # should have 1000 individuals
hist(a_freq_subset)

# vcf_full@fix[,"POS"]

rownames(G_full_subset)=vcf_full@fix[,"POS"]
colnames(G_full_subset)=indPhen_df$indID[indPhen_df$subset]

## This is my subset of individuals
dim(G_full_subset)

G_full_subset[1:20,1:10]
```

```{r set up full mutations table}
muts_full0 <- data.frame(seed = as.character(seed),
  VCFrow = 1:nrow(vcf_full@gt),
                        mutID = vcf_full@fix[,"ALT"],
                        pos_pyslim = as.integer(vcf_full@fix[,"POS"]),
                        a_freq_full = a_freq_full,
                        a_freq_subset = a_freq_subset)
dim(muts_full0)

muts_full <- merge(muts_full0, muts_df, by=c("mutID", "seed"), all.x=TRUE)
rm(muts_full0)
rm(muts_df)
dim(muts_full)

head(muts_full)
head(muts_full[muts_full$mutID!=1,]) # causal mutations

muts_full$mutTempEffect <- as.numeric(muts_full$mutTempEffect)
muts_full$mutSalEffect <- as.numeric(muts_full$mutSalEffect)


hist(muts_full$mutTempEffect, xlim=c(-1,1), xlab="Mutation temp effect", breaks=seq(-1,1,0.01), main=plotmain)
hist(muts_full$mutSalEffect, xlim=c(-1,1), xlab="Mutation Env2 effect", breaks=seq(-1,1,0.01), main=plotmain)

muts_full <- muts_full[order(muts_full$pos_pyslim),]
#plot(muts_full$VCFrow, muts_full$pos_pyslim) #sanity check

muts_full$causal <- FALSE
muts_full$causal[muts_full$mutID!=1] <- TRUE

num_causal <- sum(muts_full$causal)
num_neut <- sum(!muts_full$causal)
```


## This code estimates the correlation between AF and environment for each mutation
Used for calculation of the correlation and degree of cline
```{r af as function of env}
dim(G_full_subset)

subset_indPhen_df <- indPhen_df[indPhen_df$subset,]
rm(indPhen_df)
str(subset_indPhen_df)

head(pop_df)

subpop_subset <- unique(subset_indPhen_df$subpop)

af_pop <- matrix(NA, nrow=length(subpop_subset), ncol=nrow(G_full_subset))
rownames(af_pop) <- subpop_subset
colnames(af_pop) <- rownames(G_full_subset)

n_ind <- table(subset_indPhen_df$subpop)

subset_temp_opt <- tapply(subset_indPhen_df$temp_opt,
                          subset_indPhen_df$subpop, mean)

subset_sal_opt <- tapply(subset_indPhen_df$sal_opt,
                          subset_indPhen_df$subpop, mean)

muts_full$af_cor_temp <- NA
muts_full$af_slope_temp <- NA
muts_full$af_cor_temp_P <- NA
muts_full$af_cor_sal <- NA
muts_full$af_slope_sal <- NA
muts_full$af_cor_sal_P <- NA

for (row in 1:nrow(G_full_subset)){
    counts <- table(G_full_subset[row,], subset_indPhen_df$subpop)
      # this give a table of the number of alleles for individuals at that temperature
    
    forSum <- counts*as.numeric(rownames(counts))
      # 0 for homozygote reference
      # 1 for heterozygote
      # 2 for homozygote derived
      # by multipling counts by number, heterozygotes are counted once
      # and homozygote derived are counted twice
      # also accounts for situations when only 2 genotypes found at a locus
  af_pop[,row] <- (colSums(forSum))/(2*n_ind)
  
  # temp cors
  corrtemp <- cor.test(af_pop[,row], subset_temp_opt, method="kendall")
  muts_full$af_cor_temp[row] <- corrtemp$estimate
  muts_full$af_cor_temp_P[row] <- corrtemp$p.value
  muts_full$af_slope_temp[row] <- lm(af_pop[,row]~subset_temp_opt)$coef[2]
  
  ## do it for salinity
  corrsal <- cor.test(af_pop[,row], subset_sal_opt, method="kendall")
  muts_full$af_cor_sal[row] <- corrsal$estimate
  muts_full$af_cor_sal_P[row] <- corrsal$p.value
  muts_full$af_slope_sal[row] <- lm(af_pop[,row]~subset_sal_opt)$coef[2]
  
}

# Sanity check that what I calculate here (from the subsample)
# is similar to what I outputted from SliM in muts_df

plot( muts_full$cor_temp, muts_full$af_cor_temp, ylim=c(-1,1), xlim=c(-1,1), xlab="SliM cor(p, Temp)", ylab="Subsample cor(p, Temp)", main=plotmain)
abline(0,1)

plot(muts_full$cor_sal, muts_full$af_cor_sal, ylim=c(-1,1), xlim=c(-1,1), xlab="SliM cor(p, Env2)", ylab="Subsample cor(p, Env2)", main=plotmain)
abline(0,1)



muts_full$causal_temp <- "neutral-linked"
muts_full$causal_temp[muts_full$causal & muts_full$mutTempEffect != 0] <- "causal"
muts_full$causal_temp[muts_full$pos_pyslim > 500000] <- "neutral"

ggplot(muts_full, aes(af_cor_temp, fill=causal_temp)) + 
  geom_density(alpha=0.5) + xlab("Cor(p, temp)") + xlim(-1,1) + ggtitle(plotmain) 


muts_full$causal_sal <- "neutral-linked"
muts_full$causal_sal[muts_full$causal & muts_full$mutSalEffect != 0 & info$Ntraits == 2] <- "causal"
muts_full$causal_sal[muts_full$pos_pyslim > 500000] <- "neutral"

ggplot(muts_full, aes(af_slope_temp, fill=causal_sal)) + 
  geom_density(alpha=0.5) + xlab("Slope(p, temp)") + xlim(-1,1) + ggtitle(plotmain)



muts_full$pos_pyslim <- as.integer(muts_full$pos_pyslim)

for (LG in 1:20){
  sites_low <- (LG-1)*50000+1
  sites_high <- (LG)*50000
  muts_full$LG[muts_full$pos_pyslim >= sites_low & muts_full$pos_pyslim <= sites_high] <- LG
}


muts_full$colors <- NA
muts_full$colors[muts_full$LG %in% seq(1,9, by=2)] <- adjustcolor("goldenrod1", 0.1)
muts_full$colors[muts_full$LG %in% seq(2,10, by=2)] <- adjustcolor("goldenrod3", 0.1)
muts_full$colors[muts_full$LG %in% seq(11,19, by=2)] <- adjustcolor("grey50", 0.1)
muts_full$colors[muts_full$LG %in% seq(12,20, by=2)] <- adjustcolor("grey70", 0.1)
muts_full$colors <- as.character(muts_full$colors)

  plot(muts_full$pos_pyslim, abs(muts_full$af_cor_temp), xlab="Position", ylab="Subsample Cor(p, temp)", ylim=c(0,1), pch=19, col=muts_full$colors, main=plotmain)
  points(muts_full$pos_pyslim[muts_full$causal_temp=="causal"], abs(muts_full$af_cor_temp[muts_full$causal_temp=="causal"]), col="blue", pch=5, cex=2)
  
  plot(muts_full$pos_pyslim, abs(muts_full$af_cor_sal), xlab="Position", ylab="Subsample Cor(p, Env2)", ylim=c(0,1), pch=19, col=muts_full$colors, main=plotmain)
  points(muts_full$pos_pyslim[muts_full$causal_sal=="causal"], abs(muts_full$af_cor_temp[muts_full$causal_sal=="causal"]), col="blue", pch=5, cex=2)
  

```


This code looks at VA across metapopulation.

proportion of VA explained by locus in metapopultion (this is a little misleading because some loci are low here, but explain a lot of VA in a specific environment)
  
```{r prop Va}

muts_full$Va_temp <- muts_full$a_freq_subset*(1-muts_full$a_freq_subset)*muts_full$mutTempEffect^2

muts_full$Va_temp_prop <- muts_full$Va_temp/sum(muts_full$Va_temp, na.rm=TRUE)

hist(muts_full$Va_temp_prop, xlab="VA temp prop", main=plotmain, xlim=c(0,0.5), breaks=seq(0,1,0.01))
muts_full$Va_temp_prop[is.na(muts_full$Va_temp_prop)] <- 0
muts_full$Va_sal <- muts_full$a_freq_subset*(1-muts_full$a_freq_subset)*muts_full$mutSalEffect^2

muts_full$Va_sal_prop <- muts_full$Va_sal/sum(muts_full$Va_sal, na.rm=TRUE)
muts_full$Va_sal_prop[is.na(muts_full$Va_sal_prop)] <- 0
hist(muts_full$Va_sal_prop, xlab= "VA Env2 prop.", main= plotmain, xlim=c(0,0.5), breaks=seq(0,1,0.01))

plot(muts_full$Va_temp_prop,
     muts_full$Va_sal_prop,
     xlim=c(-0.01,1), ylim=c(-0.01,1), bty="l", pch=19, 
     xlab="Subset VA temp prop.",
     ylab="Subset VA Env2 prop.", main=plotmain)


p<- ggplot(aes(x=Va_temp_prop,y=abs(af_cor_temp), colour=causal_temp), data=muts_full) + 
  geom_point() +
 ggtitle(plotmain) + 
  theme_classic() + 
  scale_x_continuous(limits=c(0,1)) +
   scale_y_continuous(limits=c(0,1))
ggExtra::ggMarginal(p, type="violin",groupColour = TRUE, groupFill = TRUE)
   

p<- ggplot(aes(x=Va_sal_prop,y=abs(af_cor_sal), colour=causal_sal), data=muts_full) + 
  geom_point() +
 ggtitle(plotmain) + 
  theme_classic() + 
  scale_x_continuous(limits=c(0,1)) +
   scale_y_continuous(limits=c(0,1))
ggExtra::ggMarginal(p, type="violin",groupColour = TRUE, groupFill = TRUE)
     

dev.off()
```


## PCA
```{r pca}
lfmmfile <- paste0(path, seed, "_genotypes.lfmm")
write.lfmm(t(G_full_subset), lfmmfile)

pc = pca(lfmmfile, 30, scale = TRUE)

subset_indPhen_df$PC1 <- pc$projections[,1]
subset_indPhen_df$PC2 <- pc$projections[,2]
subset_indPhen_df$PC3 <- pc$projections[,3]
tw = tracy.widom(pc)
plot(tw$percentage)
a <- tw$percentage[1:20]
b <- tw$percentage[2:21]
K <- max(which(a > b*1.5))
K
```

```{r}
pdf(paste0(path,seed,"_pdf_3pca.pdf"), width=5, height=5)

plot(pc$sdev[1:10]/sum(pc$sdev), bty="l", ylab="Prop Var of PC axis", main=paste0(plotmain, "; K=", K))

ggplot(subset_indPhen_df) + geom_point(aes(x=PC1, y=PC2,size=sal_opt)) + geom_point(aes(x=PC1, y=PC2, color=temp_opt)) + theme_classic() + scale_colour_gradient2(high=rgb(1,0.4,0.2), low="cornflowerblue", mid=rgb(0.8,0.8,0.7), name="Temp") + labs(size="salinity") + ggtitle(paste0(plotmain, "; K=", K))

ggplot(subset_indPhen_df) + geom_point(aes(x=PC1, y=PC3,size=sal_opt)) + geom_point(aes(x=PC1, y=PC3, color=temp_opt)) + theme_classic() + scale_colour_gradient2(high=rgb(1,0.4,0.2), low="cornflowerblue", mid=rgb(0.8,0.8,0.7), name="Temp") + labs(size="salinity") + ggtitle(paste0(plotmain,  "; K=", K))


dev.off()
```

## LFMM
```{r lfmm2 temp}
mod_temp <- lfmm2(input = t(G_full_subset), env = subset_indPhen_df$temp_opt, K = K)
pv_temp <- lfmm2.test(object = mod_temp, input = t(G_full_subset),
                 env = subset_indPhen_df$temp_opt,
                 linear = TRUE)

muts_full$LEA3.2_lfmm2_mlog10P_tempenv <- -log10(as.numeric(pv_temp$pvalues))
muts_full$LEA3.2_lfmm2_mlog10P_tempenv_sig <- qvalue(as.numeric(pv_temp$pvalues))$qvalue < 0.05
```


```{r lfmm2 sal}
mod_sal <- lfmm2(input = t(G_full_subset), env = subset_indPhen_df$sal_opt, K = K)

pv_sal <- lfmm2.test(object = mod_sal, input = t(G_full_subset),
                 env = subset_indPhen_df$sal_opt,
                 linear = TRUE)


muts_full$LEA3.2_lfmm2_mlog10P_salenv <- -log10(as.numeric(pv_sal$pvalues))
muts_full$LEA3.2_lfmm2_mlog10P_salenv_sig <- qvalue(as.numeric(pv_sal$pvalues))$qvalue < 0.05
```

```{r}

LEA3.2_lfmm2_Va_temp_prop <- sum(muts_full$Va_temp[which(muts_full$LEA3.2_lfmm2_mlog10P_tempenv_sig)], na.rm=TRUE)/sum(muts_full$Va_temp, na.rm=TRUE) 
LEA3.2_lfmm2_Va_sal_prop <- sum(muts_full$Va_sal[which(muts_full$LEA3.2_lfmm2_mlog10P_salenv_sig)], na.rm=TRUE)/sum(muts_full$Va_sal, na.rm=TRUE) 

# True positive rate -proportion of causal loci that are outliers
LEA3.2_lfmm2_TPR_temp <- sum(muts_full$LEA3.2_lfmm2_mlog10P_tempenv_sig & muts_full$causal_temp)/sum(muts_full$causal_temp)

LEA3.2_lfmm2_TPR_sal <- sum(muts_full$LEA3.2_lfmm2_mlog10P_salenv_sig & muts_full$causal_sal)/sum(muts_full$causal_sal)

# False discovery rate- proportion of outliers that are neutral
LEA3.2_lfmm2_FDR_temp <-(sum(!muts_full$causal_temp & muts_full$LEA3.2_lfmm2_mlog10P_tempenv_sig, na.rm=TRUE)/sum(muts_full$LEA3.2_lfmm2_mlog10P_tempenv_sig, na.rm=TRUE))

LEA3.2_lfmm2_FDR_sal <- (sum(!muts_full$causal_sal & muts_full$LEA3.2_lfmm2_mlog10P_salenv_sig, na.rm=TRUE)/sum(muts_full$LEA3.2_lfmm2_mlog10P_salenv_sig, na.rm=TRUE))
```

```{r plot LFMM outliers}

pdf(paste0(path,seed,"_pdf_4manhattan_LFMM.pdf"), width=10, height=5)

  a <- ggplot() + 
    geom_point(data=muts_full, aes(x=pos_pyslim, y = LEA3.2_lfmm2_mlog10P_tempenv), color=muts_full$colors) +
    geom_point(data = muts_full[muts_full$causal_temp=="causal",], aes(x=pos_pyslim, y = LEA3.2_lfmm2_mlog10P_tempenv, 
                                                                       color=Va_temp_prop, size=Va_temp_prop), shape=17) + 
    scale_colour_viridis(option="mako", begin = 0.8, end=0, limits=c(0,1)) + 
    theme_classic() +
    geom_point(data = muts_full[muts_full$LEA3.2_lfmm2_mlog10P_tempenv_sig,], aes(x=pos_pyslim, y =LEA3.2_lfmm2_mlog10P_tempenv),pch=23, col="blue", size=3) + ylim(0, 10)  + ylab("-log10(P) LFMM temp") +
    ggtitle(paste0(plotmain," temp")) + xlab("position") + labs(color="temp VA prop.", size="temp VA prop.")
  
  b <- ggplot() + 
    geom_point(data=muts_full, aes(x=pos_pyslim, y = LEA3.2_lfmm2_mlog10P_salenv), color=muts_full$colors) +
    geom_point(data = muts_full[muts_full$causal_sal=="causal",], aes(x=pos_pyslim, y = LEA3.2_lfmm2_mlog10P_salenv, 
                                                                      color=Va_sal_prop, size=Va_sal_prop)) + 
    scale_colour_viridis(option="mako", begin = 0.8, end=0, limits=c(0,1)) + 
    theme_classic() +
    geom_point(data = muts_full[muts_full$LEA3.2_lfmm2_mlog10P_salenv_sig,], aes(x=pos_pyslim, y =LEA3.2_lfmm2_mlog10P_salenv),pch=23, col="blue", size=3) + ylim(0, 10)  + ylab("-log10(P) LFMM salinity") + ggtitle(paste0(plotmain," Env2")) + xlab("position") + labs(color="Env2 VA prop.", size="Env2 VA prop.")
  
  
grid.arrange(a, b, nrow=2)

dev.off()
```


## RDA

```{r vegan}
sal <- subset_indPhen_df$sal_opt
temp <- subset_indPhen_df$temp_opt
rdaout <- rda(t(G_full_subset)~ sal + temp)

str(rdaout)
scores <- scores(rdaout, choices=1:4)
loci.sc <- scores$species
ind.sc <- scores$sites

subset_indPhen_df$RDA1 <- ind.sc[,1]
subset_indPhen_df$RDA2 <- ind.sc[,2]

muts_full$RDA1_score <- loci.sc[,1]
muts_full$RDA2_score <- loci.sc[,2]
```

```{r rdaadapt function for outliers}
rdadapt<-function(rda,K) {
  zscores<-rda$CCA$v[,1:as.numeric(K)]
  resscale <- apply(zscores, 2, scale)
  resmaha <- covRob(resscale, distance = TRUE, na.action= na.omit, estim="pairwiseGK")$dis 
  lambda <- median(resmaha)/qchisq(0.5,df=K)
  reschi2test <- pchisq(resmaha/lambda,K,lower.tail=FALSE)
  qval <- qvalue(reschi2test)
  q.values_rdadapt<-qval$qvalues
  return(data.frame(p.values=reschi2test, q.values=q.values_rdadapt))
}
```


```{r ID RDA outliers}

ps <- rdadapt(rdaout, 2)

muts_full$RDA_mlog10P <- -log10(ps$p.values)
muts_full$RDA_mlog10P_sig <- ps$q.values<0.001



# Proportion VA
(RDA_Va_temp_prop <- sum(muts_full$Va_temp_prop[muts_full$RDA_mlog10P_sig], na.rm=TRUE))

(RDA_Va_sal_prop <- sum(muts_full$Va_sal_prop[muts_full$RDA_mlog10P_sig], na.rm=TRUE))

(RDA_TPR <- sum(muts_full$RDA_mlog10P_sig & muts_full$causal)/sum(muts_full$causal))

(RDA_FDR <- sum(muts_full$RDA_mlog10P_sig & !muts_full$causal)/sum(muts_full$RDA_mlog10P_sig))
```



```{r plot RDA outliers}

muts_scale <- c(max(abs(muts_full$RDA1_score))*1.3, max(abs(muts_full$RDA2_score))*1.3)
# for nicer plot
sal_arrow_muts = rdaout$CCA$biplot[which(rownames(rdaout$CCA$biplot)=="sal"),]*muts_scale
temp_arrow_muts = rdaout$CCA$biplot[which(rownames(rdaout$CCA$biplot)=="temp"),]*muts_scale

arrows <- data.frame(x=rep(0,2), y = rep(0, 2),
                     dx = c(sal_arrow_muts[1], temp_arrow_muts[1]),
                    dy = c(sal_arrow_muts[2], temp_arrow_muts[2]), name = c("Sal", "Temp"))

pdf(paste0(path,seed,"_pdf_5RDA.pdf"), width=5, height=5)

a<- screeplot(rdaout)
str(a)
a$y # save this it's the eigenvalues

ggplot() + 
  geom_point(data=muts_full, aes(x=RDA1_score, y = RDA2_score), color=adjustcolor("grey80", 0.2)) +
  geom_point(data = muts_full[muts_full$causal_temp=="causal",], aes(x=RDA1_score, y = RDA2_score, color=Va_temp_prop, size=Va_temp_prop), shape=17) +
  scale_colour_viridis(option="mako", begin = 0.9, end=0.2, limits=c(0,1)) +  theme_classic() +  
  geom_point(data = muts_full[muts_full$RDA_mlog10P_sig,], aes(x=RDA1_score, y = RDA2_score),pch=23, col="orange", size=3) +  xlab("RDA 1") + ylab("RDA 2") + ggtitle(paste0(plotmain, " mutations - temp")) + geom_segment(data=arrows, aes(x=x, y=y, xend=dx, yend=dy),arrow=arrow(length = unit(0.2,"cm"))) + geom_text(data=arrows, aes(x=dx, y=dy, label=name), hjust="right", vjust="bottom")+ labs(color="temp VA prop.", size="temp VA prop.")
                                   
ggplot() + 
  geom_point(data=muts_full, aes(x=RDA1_score, y = RDA2_score), color=rgb(0.4,0.4,0.6,0.1)) +
  geom_point(data = muts_full[muts_full$causal_sal=="causal",], aes(x=RDA1_score, y = RDA2_score, color=Va_sal_prop, size=Va_sal_prop)) +scale_colour_viridis(option="plasma", begin = 0.8, end=0, limits=c(0,0.5)) +  theme_classic() +  geom_point(data = muts_full[muts_full$RDA_mlog10P_sig,], aes(x=RDA1_score, y = RDA2_score),pch=23, col="grey70", size=3) +  xlab("RDA 1") + ylab("RDA 2") + ggtitle(paste0(plotmain, " mutations - Env2")) +  geom_segment(data=arrows, aes(x=x, y=y, xend=dx, yend=dy),arrow=arrow(length = unit(0.2,"cm"))) + geom_text(data=arrows, aes(x=dx, y=dy, label=name), hjust="right", vjust="bottom") + labs(color="Env2 VA prop.", size="Env2 VA prop.")
```

```{r RDA ind}
ind_scale <- c(max(abs(subset_indPhen_df$RDA1))*1.3, max(abs(subset_indPhen_df$RDA2))*1.3)
# for nicer plot
sal_arrow_muts = rdaout$CCA$biplot[which(rownames(rdaout$CCA$biplot)=="sal"),]*ind_scale
temp_arrow_muts = rdaout$CCA$biplot[which(rownames(rdaout$CCA$biplot)=="temp"),]*ind_scale

arrows_ind <- data.frame(x=rep(0,2), y = rep(0, 2),
                     dx = c(sal_arrow_muts[1], temp_arrow_muts[1]),
                    dy = c(sal_arrow_muts[2], temp_arrow_muts[2]), name = c("Env2", "Temp"))

ggplot(subset_indPhen_df) + 
  geom_point(aes(x=RDA1, y = RDA2, size=sal_opt)) + geom_point(aes(x=RDA1, y=RDA2, color=temp_opt)) + theme_classic() + scale_colour_gradient2(high=rgb(1,0.4,0.2), low="cornflowerblue", mid=rgb(0.8,0.8,0.7), name="Temp") + labs(size="Env2") + ggtitle(paste0(plotmain,"; Individs.; N traits = ", info$Ntraits)) + geom_segment(data=arrows_ind, aes(x=x, y=y, xend=dx, yend=dy),arrow=arrow(length = unit(0.2,"cm"))) + geom_text(data=arrows_ind, aes(x=dx, y=dy, label=name), hjust="right", vjust="bottom")

dev.off()
```

x <- 0:9
y <- 0:9
a <- expand.grid(x,y)
b <- round(dmvnorm(a, mean=c(4.5,4.5), sigma=matrix(c(4.98,0,0,4.98), nrow=2)),3)*10000 + 5
sum(b)
z <- cbind(a,b)
plot(z$Var1, z$Var2, col=NULL)
text(z$Var1, z$Var2, z$b)

```{r RDA manhattan}

pdf(paste0(path,seed,"_pdf_6manhattan_RDA.pdf"), width=10, height=5)

a<- ggplot() + 
  geom_point(data=muts_full, aes(x=pos_pyslim, y = RDA_mlog10P), color=rgb(0.4,0.4,0.6,0.1)) +
  geom_point(data = muts_full[muts_full$causal_temp=="causal",], aes(x=pos_pyslim, y = RDA_mlog10P, color=Va_temp_prop, size=Va_temp_prop)) + 
  scale_colour_viridis(option="plasma", begin = 0.8, end=0, limits=c(0,0.5)) + 
  theme_classic() +
  geom_point(data = muts_full[muts_full$RDA_mlog10P_sig,], aes(x=pos_pyslim, y =RDA_mlog10P),pch=23, col="grey70", size=3) + ylim(0, 10)  + ylab("-log10(P) RDA") + ggtitle(paste0(plotmain," temp")) + xlab("position") + labs(color="Temp VA prop.", size="Temp VA prop.")

b<- ggplot() + 
  geom_point(data=muts_full, aes(x=pos_pyslim, y = RDA_mlog10P), color=rgb(0.4,0.4,0.6,0.1))+
  geom_point(data = muts_full[muts_full$causal_sal=="causal",], aes(x=pos_pyslim, y = RDA_mlog10P, color=Va_sal_prop, size=Va_sal_prop)) + 
  scale_colour_viridis(option="plasma", begin = 0.8, end=0, limits=c(0,0.5)) + 
  theme_classic() +
  geom_point(data = muts_full[muts_full$RDA_mlog10P_sig,], aes(x=pos_pyslim, y =RDA_mlog10P),pch=23, col="grey70", size=3) + ylim(0, 10)  + ylab("-log10(P) RDA") + ggtitle(paste0(plotmain," sal")) + xlab("position") + labs(color="Env2 VA prop.", size="Env2 VA prop.")

grid.arrange(a, b, nrow=2)
dev.off()
```


```{r}
write.table(data.frame(seed=as.character(seed), summary(rdaout)$biplot),paste0(path,seed,"_RDAloadings.txt")) 
```


```{r predict phen as a function of random loci}
nloci <- c(10, 50, 100, 500, 50000)
Random_cor_salpredict_salphen <- rep(NA, length(nloci))
Random_cor_temppredict_tempphen <- rep(NA, length(nloci))

for (i in 1:length(nloci)){
  G_random_out <- G_full_subset[sample(1:nrow(muts_full), nloci[i], replace=FALSE),]
  rdaout_rand <- rda(t(G_random_out)~ subset_indPhen_df$sal_opt + subset_indPhen_df$temp_opt)
  scores <- scores(rdaout_rand, choices=1:4)
  loci.sc <- scores$species
  ind.sc <- scores$sites
  temp_pred <- ind.sc[,1]*eigenvals(rdaout_rand)[1]*summary(rdaout_rand)$biplot[2,1] + ind.sc[,2]*eigenvals(rdaout_rand)[2]*summary(rdaout_rand)$biplot[2,2]
  sal_pred <- ind.sc[,1]*eigenvals(rdaout_rand)[1]*summary(rdaout_rand)$biplot[1,1] + ind.sc[,2]*eigenvals(rdaout_rand)[2]*summary(rdaout_rand)$biplot[1,2]
  
  Random_cor_temppredict_tempphen[i] <- cor(scale(subset_indPhen_df$phen_temp), scale(temp_pred), method = "kendall")
  Random_cor_salpredict_salphen[i] <- cor(scale(subset_indPhen_df$phen_sal), scale(sal_pred), method = "kendall")
}

pdf(paste0(path,seed,"_pdf_6zRDA_predict.pdf"), width=10, height=5)

plot(nloci, Random_cor_temppredict_tempphen, type="l", ylim=c(-0.1,1), ylab="Cor(RDA prediction, true phenotype)", bty="l" , xlab="Number random loci in RDA", col="darkred", lwd=2, xlim=c(0, 50000), main = paste0(plotmain), cex.main=0.5, las=2)
  points(nloci, Random_cor_salpredict_salphen, type="l", col="cornflowerblue", lwd=2, lty=2)
  legend(10000, 0.1, c("Temp", "Env2"), col=c("darkred", "cornflowerblue"), lty = c(1,2), bty="n")
  
dev.off()

out_RDA_pred <- data.frame(nloci, Random_cor_temppredict_tempphen, Random_cor_salpredict_salphen)

cor_RDA50000_RDloadings_tempPhen <- Random_cor_temppredict_tempphen[nloci==50000]
cor_RDA50000_RDloadings_salPhen <- Random_cor_salpredict_salphen[nloci==50000]

write.table(out_RDA_pred,paste0(path,seed,"_RDA_predictions.txt"))
```


## This code estimates the AF of all populations at the same temp/sal
Used for plotting
```{r af as function of salinity}

str(af_pop)
subset_temp_opt
subset_sal_opt

sal_levels <- sort(as.numeric(unique(subset_indPhen_df$sal_opt)))
sal_levels
temp_levels <- sort(as.numeric(unique(subset_indPhen_df$temp_opt)))
temp_levels

af_sal <- matrix(NA, nrow=length(sal_levels), ncol=nrow(G_full_subset))
af_temp <- matrix(NA, nrow=length(temp_levels), ncol=nrow(G_full_subset))

# WARNING: this code only works with equal sample sizes for each population
for (i in 1:ncol(af_pop)){
  af_temp[,i] <- tapply(af_pop[,i], subset_temp_opt, mean)
  af_sal[,i] <- tapply(af_pop[,i], subset_sal_opt, mean)
}
colnames(af_temp) <- muts_full$mutID
v <- tapply(af_pop[,1], subset_temp_opt, mean)
rownames(af_temp) <- as.numeric(dimnames(v)[[1]])

colnames(af_sal) <- muts_full$mutID
v <- tapply(af_pop[,1], subset_sal_opt, mean)
rownames(af_sal) <- as.numeric(dimnames(v)[[1]])


muts_full$af_cor_temp_pooled <- as.numeric(cor(af_temp, temp_levels))
muts_full$af_cor_sal_pooled <- as.numeric(cor(af_sal, sal_levels))

pdf(paste0(path,seed,"_pdf_7_afcors.pdf"), width=5, height=5)

ggplot() + 
  geom_point(data=muts_full[muts_full$causal_temp=="neutral",], aes(x=af_cor_temp, y = af_cor_temp_pooled), color=adjustcolor("grey", 0.1)) +
  geom_point(data = muts_full[muts_full$causal_temp=="causal",], aes(x=af_cor_temp, y = af_cor_temp_pooled,  size=Va_temp_prop, color=Va_temp_prop), alpha=0.8) +
  scale_colour_viridis(option="plasma", begin = 1, end=0.2, limits=c(0,0.5)) +  theme_classic() +  xlab("Cor(af, temp)") + 
  ylab("Cor(af, temp) pooled by temp.") + ggtitle(paste0(plotmain, " mutations: temp")) + ylim(-1,1) + xlim(-1, 1) + geom_abline(data=NULL, intercept=0, slope=1) + labs(color="temp VA prop.", size="temp VA prop.")

ggplot() + 
  geom_point(data=muts_full, aes(x=af_cor_sal, y = af_cor_sal_pooled), color=rgb(0.4,0.4,0.6,0.1)) +
  geom_point(data = muts_full[muts_full$causal_sal=="causal",], aes(x=af_cor_temp, y = af_cor_sal_pooled, color=Va_sal_prop, size=Va_sal_prop), alpha=0.8) +scale_colour_viridis(option="plasma", begin = 0.8, end=0, limits=c(0,0.5)) +  theme_classic() +  xlab("Cor(af, Env2.)") + ylab("Cor(af, Env2) pooled by Env2.") + ggtitle(paste0(plotmain, " mutations: Env2")) + ylim(-1,1) + xlim(-1, 1) + geom_abline(data=NULL, intercept=0, slope=1) + labs(color="Env2 VA prop.", size="Env2 VA prop.")

dev.off()
```


# Plot phenotype vs. AF clines

1st column:
Phenotypic clines for temp (top) and salinity (bottom)

2nd column:
AF clines for temp (top) and salinity (bottom)

```{r}

pdf(paste0(path,seed,"_pdf_8phen-env_af-env.pdf"), width=9, height=9)

  par(mfrow=c(2,2), mar=c(4,4,1,1), oma=c(0,0,2,1))


  plot(jitter(subset_indPhen_df$temp_opt), subset_indPhen_df$phen_temp, xlab="Deme Temperature", ylab="Ind. optimum temp.",pch=18, , las=1, bty="n", col=adjustcolor("purple",0.1))
  abline(lm(subset_indPhen_df$phen_temp~subset_indPhen_df$temp_opt))
    
  plot(jitter(subset_indPhen_df$sal_opt), subset_indPhen_df$phen_sal,
       xlab="Deme Salinity", ylab="Ind. optimum sal.", pch=20, col=adjustcolor("blue",0.1), las=1, bty="n"
       )
  abline(lm(subset_indPhen_df$phen_sal~subset_indPhen_df$sal_opt))
  
mtext(paste0(plotmain), side=3, outer=TRUE, cex=0.5, line=1)



#viridis(10)
#col <- two.colors(10, "#3B528BFF", "#FDE725FF", middle = "#6DCD59FF")
col <- rev(mako(10))

### TEMP PLOT ####
##################### ###
plot(pop_df$opt1, rep(0, length(pop_df$opt1)), ylim=c(0,1), xlab="Deme Temperature", ylab="Allele frequency", col=rgb(0,0,0,0), bty="n", xlim=c(-1, 1), las=1)

str(af_temp)
for (i in which(muts_full$causal_temp=="causal")){
  cor_mut <- cor(temp_levels, af_temp[,i])
  if(muts_full$Va_temp_prop[i]>0.05){
    lwd=3
    alpha1 = 0.8
  }else{
      lwd=0.5
      alpha1 = 0.5
      }
    lines(temp_levels, af_temp[,i], col=adjustcolor(col[ceiling(abs(cor_mut*10))], alpha1), lwd=lwd)
}


### SAL PLOT ####
##################### ###
  plot(pop_df$opt0, rep(0, length(pop_df$opt0)), ylim=c(0,1), xlab="Deme Env2", ylab="Allele frequency", col=rgb(0,0,0,0), bty="n", xlim=c(-1, 1), las=1)

for (i in which(muts_full$causal_sal=="causal")){
  cor_mut <- cor(sal_levels, af_sal[,i])
  if(muts_full$Va_sal_prop[i]>0.05){
    lwd=3
    alpha1 = 0.8
  }else{
      lwd=0.5
      alpha1 = 0.5
      }
    lines(sal_levels, af_sal[,i], col=adjustcolor(col[ceiling(abs(cor_mut*10))], alpha1), lwd=lwd)
}



### LEGEND PLOT ####
##################### ###
par(mfrow=c(1,1))
plot(pop_df$opt1, rep(0, length(pop_df$opt1)), ylim=c(0,1), xlab="Temperature", ylab="Allele frequency", col=rgb(0,0,0,0), bty="n", xlim=c(-1, 1), main=plotmain)

legend(0,1,seq(0.05,0.95, length.out=10), fill=col, cex=0.7, bty="n", title="Cor (p, env)")

dev.off()

```

```{r genotype heatmaps}

pdf(paste0(path,seed,"_pdf_heatmaps.pdf"), width=5, height=5)

# Temps
order_temp <- order(subset_indPhen_df$temp_opt)
order_sal <- order(subset_indPhen_df$sal_opt)

par(mfrow=c(1,1), mar=c(4,0,6,4))

heatmap(t(G_full_subset[muts_full$causal_temp=="causal",order_temp]), Rowv = NA,  main=paste0(plotmain," Causal alleles Temp"),cexCol = 0.3,  useRaster=FALSE, 
        scale="none",
           #Colv = NA,
           labRow = round(subset_indPhen_df$temp_opt[order_temp],1), xlab="Mutation ID", ylab="South <---------> North")

if(sum(muts_full$causal_sal=="causal")>0){
  heatmap(t(G_full_subset[muts_full$causal_sal=="causal",order_sal]), Rowv = NA,  main=paste0(plotmain,"Causal alleles Env2"),cexCol = 0.3,  useRaster=FALSE,
        scale="none", xlab="Mutation ID", ylab="East <---------> West",
           #Colv = NA,
           labRow = round(subset_indPhen_df$temp_opt[order_temp],1))
}


  heatmap(t(G_full_subset[sample(which(!muts_full$causal), 1000), order_temp]), Rowv = NA,  main=paste0(plotmain,"Non-causal alleles"),cexCol = 0.3,  useRaster=FALSE,
        scale="none",
           #Colv = NA,
           labRow = round(subset_indPhen_df$temp_opt[order_temp],1), xlab="Mutation ID", ylab="South <---------> North")


### Visualize genotypes at low salinities ###
subset_indPhen_df2 <- subset_indPhen_df[subset_indPhen_df$sal_opt==-1,]
order_temp2 <- order(subset_indPhen_df2$temp_opt)

G_full_subset2 <- G_full_subset[,subset_indPhen_df$sal_opt==-1]

heatmap(t(G_full_subset2[muts_full$causal,order_temp2]), Rowv = NA,  main=paste0(plotmain," Low Env2. Demes; All causal alleles"),cexCol = 0.3,  useRaster=FALSE, 
        scale="none",
           #Colv = NA,
           labRow = round(subset_indPhen_df2$temp_opt[order_temp2],1), xlab="Mutation ID", ylab="South <---------> North")

## Non-causal
heatmap(t(G_full_subset2[sample(which(!muts_full$causal), 1000),order_temp2]), Rowv = NA,  main=paste0(plotmain," Low Sal. Demes; Non-causal alleles"),cexCol = 0.3,  useRaster=FALSE, 
        scale="none",
           #Colv = NA,
           labRow = round(subset_indPhen_df2$temp_opt[order_temp2],1), xlab="Mutation ID", ylab="South <---------> North")


### Visualize genotypes at high salinities ###
subset_indPhen_df3 <- subset_indPhen_df[subset_indPhen_df$sal_opt==1,]
order_temp3 <- order(subset_indPhen_df3$temp_opt)

G_full_subset3 <- G_full_subset[,subset_indPhen_df$sal_opt==1]

heatmap(t(G_full_subset3[muts_full$causal,order_temp3]), Rowv = NA,  main="High Sal. Demes; Causal alleles",cexCol = 0.3,  useRaster=FALSE,
        scale="none",
           #Colv = NA,
           labRow = round(subset_indPhen_df2$temp_opt[order_temp3],1), xlab="Mutation ID", ylab="South <---------> North")


heatmap(t(G_full_subset3[sample(which(!muts_full$causal), 100),order_temp3]), Rowv = NA,  main="High Sal. Demes; Non-causal alleles",cexCol = 0.3,  useRaster=FALSE,
        scale="none",
           #Colv = NA,
           labRow = round(subset_indPhen_df2$temp_opt[order_temp3],1), xlab="Mutation ID", ylab="South <---------> North")

dev.off()

rm(subset_indPhen_df2)
rm(subset_indPhen_df3)
rm(G_full_subset2)
rm(G_full_subset3)
```

```{r write outputs}
    Bonf_alpha <- (0.05/(num_causal + num_neut))
    
    num_causal_sig_temp_corr <- sum(muts_full$af_cor_temp_P[muts_full$causal_temp=="causal"]<Bonf_alpha)# number of causal loci that have significant Spearman's correlations with temperature after Bonferroni correction
    
    num_causal_sig_sal_corr<- sum(muts_full$af_cor_sal_P[muts_full$causal_sal=="causal"]<Bonf_alpha)# number of causal loci that have significant Spearman's correlations with salinity after Bonferroni correction
    
    num_neut_sig_temp_corr <- sum(muts_full$af_cor_temp_P[!muts_full$causal_temp=="causal"]<Bonf_alpha)# number of neutral loci that have significant  Spearman's correlations with temperature after Bonferroni correction
    
    num_neut_sig_sal_corr <- sum(muts_full$af_cor_sal_P[!muts_full$causal_sal=="causal"]<Bonf_alpha)# proportion of neutral loci that have significant Spearman's correlations with salinity after Bonferroni correction
      
    median_causal_temp_cor <- median(abs(muts_full$af_cor_temp[muts_full$causal_temp=="causal"]))#median abs(Spearman's correlation) between allele frequency and temperature for causal loci
    median_causal_sal_cor <- median(abs(muts_full$af_cor_sal[muts_full$causal_sal=="causal"])) #median Spearman's correlation  between allele frequency and salinity for causal loci
    
    median_neut_temp_cor <- median(abs(muts_full$af_cor_temp[!muts_full$causal_temp=="causal"])) # median Spearman's correlation  between allele frequency and temperature for neutral loci
    
    median_neut_sal_cor <-  median(abs(muts_full$af_cor_sal[!muts_full$causal_sal=="causal"]))# median Spearman's correlation  between allele frequency and salinity for neutral loci 
    
    write.table(muts_full, paste0(path,seed,"_muts_full.txt"))
    write.table(subset_indPhen_df, paste0(path,seed,"_ind_subset.txt"))
    
    out_full <- data.frame(seed, n_samp=n, K,
                           all_corr_phen_temp,
                           subsamp_corr_phen_temp,
                           all_corr_phen_sal,
                           subsamp_corr_phen_sal,
                           num_causal,
                           num_neut,
                           prop_causal_sig_temp_corr,
                           LEA3.2_lfmm2_Va_temp_prop, 
                           LEA3.2_lfmm2_Va_sal_prop, 
                           LEA3.2_lfmm2_TPR_temp,
                           LEA3.2_lfmm2_TPR_sal,
                           LEA3.2_lfmm2_FDR_temp, 
                           LEA3.2_lfmm2_FDR_sal, 
                           RDA_Va_temp_prop,
                           RDA_Va_sal_prop,
                           RDA_TPR, 
                           RDA_FDR, 
                           cor_RDA50000_RDloadings_tempPhen, 
                           cor_RDA50000_RDloadings_salPhen,
                           num_causal_sig_temp_corr,
                           num_causal_sig_sal_corr,
                           num_neut_sig_temp_corr,
                           num_neut_sig_sal_corr,
                           median_causal_temp_cor,
                           median_causal_sal_cor,
                           median_neut_temp_cor,
                           median_neut_sal_cor
                           )
```


