---
title: "set up non-clinal sims"
author: "KE Lotterhos"
date: "6/7/2021"
output: html_document
---

setwd("/Users/lotterhos/Documents/GitHub/MVP-NonClinalAF")

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
runID = "20220201"
```

# 15 demographies
The three categories of demographies were designed to give insights into different evolutionary processes: “Estuary Clines” (Est-Clines), “Stepping-Stone Clines” (SS-Clines), and  “Stepping-Stone Mountain Range” (SS-Mtn,  with a latitudinal cline in temperature and non-monotonic longitudinal pattern in elevation).

```{r demogs}
(demog_level = factor(c("SS-Clines", "SS-Mtn", "Est-Clines"), levels=c("SS-Clines", "SS-Mtn", "Est-Clines"), ordered=TRUE))
MIG_x = c(0.03, 0.03, 0.49)
MIG_y = c(0.03, 0.03, 0.07)
xcline = c("linear", "V", "linear")
ycline = c("linear", "linear", "linear")
demog = c("SS", "SS", "Estuary")
METAPOP_SIDE_x = 10
METAPOP_SIDE_y = 10
(demog_base_df <- data.frame(demog_level, MIG_x, MIG_y, xcline, ycline, demog, METAPOP_SIDE_x, METAPOP_SIDE_y))
```

Within these three broad categories, there were 5 sub-categories with different patterns of population size for each deme (N_deme) or migration. Three of these had base levels of m_x and m_y as described above and N_deme was manipulated: (i) equal N, (ii) N latitudinal cline (which confounded the amount of drift with the selective environment), and (iii) N largest in range center. The total metapopulation size (N_metapopulation = 10,000) was controlled for and constant across these three scenarios. The fourth sub-category kept N constant for each deme, but introduced two biogeographic breaks along the latitudinal cline, to introduce a higher degree of confounding between population structure and adaptation in the temperature trait. The fifth sub-category, which was designed to investigate a complex case, simulated variable migration (sampled from c(0.001,0.01,0.1,0.25)) between adjacent demes and variable N (sampled from c(10,10,50,50,95,95,145,145,200,200)) for each deme. 

```{r sub-demogs}
(demog_level_sub <- factor(c("N-equal_m-constant", "N-cline-N-to-S_m-constant", "N-cline-center-to-edge_m-constant", "N-equal_m_breaks", "N-variable_m-variable"), levels=c("N-equal_m-constant", "N-cline-N-to-S_m-constant", "N-cline-center-to-edge_m-constant", "N-equal_m_breaks", "N-variable_m-variable"), ordered=TRUE))
Nequal = c(0, 2, 4, 0, 3)
isVariableM = c(0,0,0,0, 1)
MIG_breaks = c(0, 0,0,1, 0)
# 0 for equal N, 1 for N cline on x-axis, 2 for N cline on y-axis, 3 for variable N, 4 for N largest in range center
# NOTE: the variable N only works for grids of length 10

(demog_sub_df <- data.frame(demog_level_sub, Nequal, isVariableM, MIG_breaks))
```

Get the 15 demography parameters
```{r}
a <- expand.grid(demog_level=demog_level, demog_level_sub=demog_level_sub)
a

a2 <- merge(a, demog_base_df, sort=FALSE)
demog_df <- merge(a2, demog_sub_df, sort=FALSE)

demog_df <- demog_df[order(demog_df$demog_level_sub, demog_df$demog_level),]
demog_df
```

# architectures

The 10 genetic architectures were broadly divided into 3 categories: highly polygenic (many loci of small effect that affect the trait), moderately polygenic, and oligogenic (few loci of large effect that affect the trait). 

```{r arch base}
arch_level <- factor(c("oliogenic", "mod-polygenic", "highly-polygenic"), ordered=TRUE)
MU_base <- 1e-07
MU_QTL_proportion <- c(0.001, 0.1, 0.25) # proportion of mutations that are QTNs
SIGMA_QTN_1 <- c(0.4, 0.1, 0.002)
SIGMA_QTN_2 <- c(0.4, 0.1, 0.002) #standard deviation of mutational effect size - for trait 2

arch_base_df <- data.frame(arch_level, MU_base, MU_QTL_proportion, SIGMA_QTN_1, SIGMA_QTN_2)
```

Within each category were 5 sub-categories: 1 trait (which adapted to the latitudinal cline, the most commonly simulated scenario), 2 traits with pleiotropy (QTNs could evolve effects on one or both traits) and equal strengths of selection on both traits, 2 traits with pleiotropy and with weaker selection on the latitudinal trait, 2 traits without pleiotropy (QTNs effect were restricted to one trait) and equal strengths of selection on both traits, and 2 traits without pleiotropy and with weaker selection on the latitudinal trait.


```{r arch sub}
arch_level_sub <- factor(c("1-trait", "2-trait-no-pleiotropy-equal-S", "2-trait-no-pleiotropy-unequal-S", "2-trait-pleiotropy-equal-S", "2-trait-pleiotropy-unequal-S"), levels=c("1-trait", "2-trait-no-pleiotropy-equal-S", "2-trait-no-pleiotropy-unequal-S", "2-trait-pleiotropy-equal-S", "2-trait-pleiotropy-unequal-S"), ordered=TRUE)
arch_level_sub

SIGMA_K_1 <- c(0.5, 0.5, 0.5, 0.5, 0.5) # trait 1 is x-axis cline/env
SIGMA_K_2 <- c(0.5, 0.5, 4.0, 0.5, 4.0) # trait 2 is y-axis cline/env

N_traits <- c(1, 2, 2, 2, 2)

ispleiotropy <- c(0, 0, 0, 1, 1)

arch_sub_df <- data.frame(arch_level_sub, SIGMA_K_1, SIGMA_K_2, N_traits, ispleiotropy)

arch_sub_df
```


Combine to get all 15 levels of architecture
```{r}
b <- expand.grid(arch_level = arch_level, arch_level_sub=arch_level_sub)
b

b2 <- merge(b, arch_base_df, sort=FALSE)
arch_df <- merge(b2, arch_sub_df, sort=FALSE)

arch_df <- arch_df[order(arch_df$arch_level, arch_df$arch_level_sub),]
arch_df
```

Combine demographies and architectures
```{r}
arch_df$arch <- paste(arch_df$arch_level, arch_df$arch_level_sub, sep="_")
demog_df$demog_name <- paste(demog_df$demog_level, demog_df$demog_level_sub, sep="_")
all <- expand.grid(demog_name = demog_df$demog_name, arch = arch_df$arch)

dim(all)

all1 <- merge(all, demog_df)
dim(all1)
all2 <- merge(all1, arch_df)
dim(all2)

str(all2)
```

Add seed 
```{r}
all2$level <- paste(all2$arch, all2$demog_name, sep="__")
reps <- 1:10

final0 <- expand.grid(reps=reps, level=all2$level)
final <- merge(final0, all2)
final <- final[order(final$reps),]
final$seed <- 1231094:(1231094 + nrow(final) - 1)

str(final)
```


```{r}
longruns <- which(final$demog_name=="Est-Clines_N-variable_m-variable" | final$demog_name=="Est-Clines_N-equal_m_breaks")

towrite <- final[-longruns,]
write.table(towrite[1:999,], paste0("src/0b-final_params-fastruns-",runID,".txt"), row.names=FALSE)
write.table(towrite[1000:nrow(towrite),], paste0("src/0b-final_params-fastruns-",runID,"-b.txt"), row.names=FALSE)
nrow(final[longruns,])
write.table(final[longruns,], paste0("src/0b-final_params-longruns-",runID,".txt"), row.names=FALSE)

save(final, file=paste0("src/0b-final_params-",runID,".RData"))

# head(final[final$N_traits==2,], 20)
```

