---
title: "Non-clinal sims final analysis"
author: "KE Lotterhos"
date: "9/7/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(ggplot2)
library(viridis)
library(gridExtra)
library(tidyr)
```

`setwd("~/Documents/GitHub/MVP-NonClinalAF/src")`

## Read in data

```{r cars}
out.df <- read.table(file = "../summary.txt", header=TRUE)
head(out.df)
tail(out.df)

sims.df <- read.table(file="0b-final_params_20210830.txt", header=TRUE)
head(sims.df)

final.df <- merge(out.df, sims.df, all.x=TRUE)

vars <- t(final.df[1,])
```


```{r}
hist(final.df$final_LA, xlim=c(0,1), breaks=seq(0,1,0.05), main="", xlab="Degree of local adaptation")
```



## Statistical model

AUCPR ~ trait + demog_level + demog_level_sub + arch_level_sub + arch_level +  + final_LA

```{r}

stat_df <- gather(final.df, key=trait, value=cor_TPR, cor_TPR_temp, cor_TPR_sal)
tail(stat_df)

unique(stat_df$demog_level)
unique(stat_df$demog_level_sub)
unique(stat_df$demog_name)
unique(stat_df$arch_level_sub)
unique(stat_df$arch)

## A model with more hierarchy
modelsum <- summary(aov(cor_TPR ~ trait + demog_level + demog_level_sub + arch_level_sub + arch_level + final_LA + trait*demog_level + trait*arch_level_sub, data=stat_df))

modelsum

pervar <- data.frame(variable = rownames(modelsum[[1]]), SS = modelsum[[1]][,2])

pervar$pervar <- pervar$SS/sum(pervar$SS)*100

pervar

## A model with less hierarchy
modelsum <- summary(aov(cor_TPR ~ trait + demog_name + arch  + trait*demog_name + trait*arch + demog_name*arch +  final_LA, data=stat_df))
modelsum

pervar <- data.frame(variable = rownames(modelsum[[1]]), SS = modelsum[[1]][,2])

pervar$pervar <- round(pervar$SS/sum(pervar$SS)*100, 1)

pervar

```

## How does demography contribute to the evolution of phenotypic clines without clines in causal allele frequencies?

Number of loci with correlations
```{r}
boxplot(final.df$cor_TPR_temp~final.df$demog_level, ylim=c(0,1))

ggtheme <- theme_bw() + theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank(), panel.background = element_blank(), panel.border=element_blank(), axis.line = element_line(colour="grey30"), axis.title = element_text(colour="grey20"), axis.text = (element_text(colour="grey30")), legend.title = element_text(colour="grey20"), legend.text = element_text(colour="grey30"))

final.df$demog_level <- factor(final.df$demog_level, levels=c("SS-Clines", "SS-Mtn", "Est-Clines"), ordered?=TRUE)

a<- ggplot(final.df, aes(x=as.factor(N_traits), y=cor_TPR_temp, fill=demog_level)) + geom_boxplot(color="grey") + ggtheme + xlab("Number of traits") + ylab("Proportion of causal temp. \n loci with significant clines") + scale_color_viridis(discrete=TRUE) + theme(legend.position = "none") + ylim(0,1)


forb <- final.df[final.df$N_traits==2,]
forb$N_traits <- factor(forb$N_traits)
b<- ggplot(forb, aes(x=N_traits, y=cor_TPR_sal, fill=demog_level)) + geom_boxplot(color="grey") + ggtheme + xlab("Number of traits") + ylab("Proportion of causal Env2 \n loci with significant clines") + scale_color_viridis(discrete=TRUE) + ylim(0,1)


#pdf("PropCausalAllelesSigClines.pdf", width)
grid.arrange(a,b, nrow=1)
```

Percent of Va explained by clinal loci
```{r}

a1<- ggplot(final.df, aes(x=as.factor(N_traits), y=cor_VA_temp_prop, fill=demog_level)) + geom_boxplot(color="grey") + ggtheme + xlab("Number of traits") + ylab("Proportion of VA in temp. \n explained by loci with significant clines") + scale_color_viridis(discrete=TRUE) + theme(legend.position = "none")+ ylim(0,1)

b1<- ggplot(forb, aes(x=N_traits, y=cor_VA_sal_prop, fill=demog_level)) + geom_boxplot(color="grey") + ggtheme + xlab("Number of traits") + ylab("Proportion of VA in Env2 \n loci  explained by loci with significant clines") + scale_color_viridis(discrete=TRUE) + ylim(0,1)


#pdf("PropCausalAllelesSigClines.pdf", width)
grid.arrange(a1,b1, nrow=1)
```

```{r}

unique(final.df$demog)
unique(final.df$demog_level)
unique(final.df$demog_level_sub)

final.df$demog_level_sub<- factor(final.df$demog_level_sub, levels=c("N-equal_m-constant", "N-cline-N-to-S_m-constant", "N-cline-center-to-edge_m-constant", "N-equal_m_breaks", "N-variable_m-variable"))

g<- ggplot(final.df, aes(x=as.factor(demog_level_sub), y=cor_TPR_temp, fill=demog_level)) + geom_boxplot(color="grey") + ggtheme + xlab("") + ylab("Proportion of causal temp. \n loci with significant clines") + scale_color_viridis(discrete=TRUE) + theme(legend.position = "none") + theme( axis.text.x=element_blank())

h <- ggplot(final.df, aes(x=as.factor(demog_level_sub), y=cor_TPR_sal, fill=demog_level)) + geom_boxplot(color="grey") + ggtheme + xlab("Demography sublevel") + ylab("Proportion of causal Env2 \n loci with significant clines") + scale_color_viridis(discrete=TRUE) + theme(legend.position = "none") + theme( axis.text.x=element_text(angle=90,vjust=1, size = 10))

grid.arrange(g,h, nrow=7, ncol=1, layout_matrix=matrix(c(1,1,2,2,2,2,2), nrow=7))
```

## How does genetic architecture contribute to the evolution of phenotypic clines without clines in causal allele frequencies?

```{r}

c <- ggplot(final.df, aes(x=as.factor(arch_level), y=cor_TPR_temp, fill=demog_level)) + geom_boxplot(color="grey") + ggtheme + xlab("Genic level") + ylab("Proportion of causal temp. \n loci with significant clines") + scale_color_viridis(discrete=TRUE) + theme(legend.position = "none") + ylim(0,1)

d<- ggplot(final.df, aes(x=as.factor(arch_level), y=cor_TPR_sal, fill=demog_level)) + geom_boxplot(color="grey") + ggtheme + xlab("Genic level") + ylab("Proportion of causal Env2 \n loci with significant clines") + scale_color_viridis(discrete=TRUE) + theme(legend.position = "none") + ylim(0,1)

grid.arrange(a,b,c,d, nrow=2, ncol=2)

unique(final.df$arch)
unique(final.df$arch_level)
unique(final.df$arch_level_sub)
```

# Architecture sub level
```{r}
e <- ggplot(final.df, aes(x=as.factor(arch), y=cor_TPR_temp, fill=demog_level)) + geom_boxplot(color="grey") + ggtheme + xlab("Genic level") + ylab("Proportion of causal temp. \n loci with significant clines") + scale_color_viridis(discrete=TRUE) + theme(legend.position = "none") + theme( axis.text.x=element_blank())

f <- ggplot(final.df, aes(x=as.factor(arch), y=cor_TPR_sal, fill=demog_level)) + geom_boxplot(color="grey") + ggtheme + xlab("Genic level") + ylab("Proportion of causal Env2 \n loci with significant clines") + scale_color_viridis(discrete=TRUE) + theme(legend.position = "none") + theme( axis.text.x=element_text(angle=90, size = 10))

grid.arrange(e,f, nrow=7, ncol=1, layout_matrix=matrix(c(1,1,2,2,2,2,2), nrow=7))

#TO DO: seperate out equal_S from un_equal S to show WHAT IT DOES
```

### Pleiotropy vs no pleiotropy

```{r}
subdf_arch <- final.df[final.df$N_traits==2,]

e1<- ggplot(subdf_arch, aes(x=as.factor(ispleiotropy), y=cor_TPR_temp)) + geom_boxplot(color="grey", fill="navyblue") + ggtheme + xlab("") + ylab("Proportion of causal temp. \n loci with significant clines") + scale_color_viridis(discrete=TRUE) + theme(legend.position = "none") + theme( axis.text.x=element_blank())

unique(final.df$SIGMA_K_1)
unique(subdf_arch$SIGMA_K_2)
subdf_arch$SIGMA_K_2 <- as.factor(subdf_arch$SIGMA_K_2)

my_col_arch <- c(viridis(10)[3], viridis(10)[7])

f1 <-   ggplot(subdf_arch, aes(x=arch_level, y=cor_TPR_temp, fill=as.factor(ispleiotropy))) + geom_boxplot(color="grey") + ggtheme + xlab("") + ylab("Proportion of causal temp. \n loci with significant clines") + scale_fill_manual( values=my_col_arch, labels=c("No pleiotropy", "Pleiotropy"), name="2-trait architectures")  + ylim(0,1) #+ theme(legend.title=element_text("2-trait\narchitectures"))

f2 <-   ggplot(subdf_arch, aes(x=arch_level, y=cor_TPR_sal, fill=as.factor(ispleiotropy))) + geom_boxplot(color="grey") + ggtheme + xlab("") + ylab("Proportion of causal sal. \n loci with significant clines") + scale_fill_manual( values=my_col_arch, labels=c("No pleiotropy", "Pleiotropy"), name="2-trait architectures")  + ylim(0,1) #+ theme(legend.title=element_text("2-trait\narchitectures"))
f2

TO DO: Combine data together for a single plot f1/f2
#unique(subdf_arch$SIGMA_K_2)

pleiotropy.df <- gather(subdf_arch, key=trait, value=cor_TPR, cor_TPR_temp, cor_TPR_sal)

f2 <-   ggplot(pleiotropy.df, aes(x=arch_level, y=cor_TPR, fill=as.factor(ispleiotropy))) + geom_boxplot(color="grey") + ggtheme + xlab("") + ylab("Proportion of causal sal. \n loci with significant clines") + scale_fill_manual( values=my_col_arch, labels=c("No pleiotropy", "Pleiotropy"), name="2-trait architectures")  + ylim(0,1)

grid.arrange(a,b,f2, nrow=2, ncol=2, layout_matrix=matrix(c(1,2,3,3), nrow=2, byrow=TRUE))
```

Strength of selection
```{r}
my_col_S <- c(mako(10)[2], mako(10)[8])

ggplot(subdf_arch, aes(x=arch_level, y=cor_TPR_temp, fill=as.factor(SIGMA_K_2))) + geom_boxplot(color="grey") + ggtheme + xlab("") + ylab("Proportion of causal temp. \n loci with significant clines") + scale_fill_manual( values=my_col_S, labels=c("Strong selection (sigma_K=0.5)", "Weak selection (sigma_K=4)"), name="2-trait architectures")  + ylim(0,1) #+ theme(legend.title=element_text("2-trait\narchitectures"))

ggplot(subdf_arch, aes(x=arch_level, y=cor_TPR_sal, fill=as.factor(SIGMA_K_2))) + geom_boxplot(color="grey") + ggtheme + xlab("") + ylab("Proportion of causal sal. \n loci with significant clines") + scale_fill_manual( values=my_col_S, labels=c("Strong selection on other trait (sigma_K=0.5)", "Weak selection on other trait (sigma_K=4)"), name="2-trait architectures")  + ylim(0,1) #+ theme(legend.title=element_text("2-trait\narchitectures"))
```

## Under what conditions can we reliably identify the alleles with effects on adaptive phenotypes using genetic-environment associations (latent factor mixed models, uncorrected correlations, and redundancy analysis)? 

```{r}

comparemethods.AUC.temp <- gather(final.df, key=method, value=AUCPR, cor_AUCPR_temp_neutSNPs, LEA3.2_lfmm2_AUCPR_temp_neutSNPs, RDA_AUCPR_neutSNPs)
comparemethods.AUC.sal <- gather(final.df, key=method, value=AUCPR, cor_AUCPR_sal_neutSNPs, LEA3.2_lfmm2_AUCPR_sal_neutSNPs, RDA_AUCPR_neutSNPs)

whichcols <- c("method", "AUCPR", "cor_PC1_temp", "cor_PC1_sal","cor_LFMMU1_temp", "cor_LFMMU1_sal", "cor_LFMMU2_temp", "cor_LFMMU2_sal")

comparemethods.AUC <- rbind(comparemethods.AUC.temp[,whichcols], comparemethods.AUC.sal[whichcols])

g<- ggplot(comparemethods.AUC, aes(x=as.factor(method), y=AUCPR, fill=method)) + geom_boxplot(color="grey") + ggtheme + xlab("method") + ylab("AUC-PR") + scale_fill_viridis(option="mako", discrete=TRUE) + theme(legend.position = "none")


# Does structure
unique(final.df$arch_level)

final.df$maxcor_LFMMU_temp <- NA
for (i in 1:nrow(final.df)){
  final.df$maxcor_LFMMU_temp[i] <- max(abs(final.df$cor_LFMMU1_temp[i]), abs(final.df$cor_LFMMU2_temp[i]))
}


h <- ggplot(final.df, aes(x=abs(cor_LFMMU1_temp), y=LEA3.2_lfmm2_AUCPR_temp_neutSNPs, color=as.factor(N_traits))) + geom_point() + ggtheme + scale_color_viridis(option="mako", begin=.20, end=.80, discrete=TRUE) +  ylim(0,1) + ylab("AUC-PR temp") + geom_smooth() + xlim(0,1)

i <- ggplot(final.df, aes(x=abs(cor_LFMMU1_sal), y=LEA3.2_lfmm2_AUCPR_sal_neutSNPs, color=as.factor(N_traits))) + geom_point() + ggtheme + scale_color_viridis(option="mako", begin=.20, end=.80, discrete=TRUE) + geom_smooth() + ylim(0,1)+ ylab("AUC-PR Env2")+ xlim(0,1)

#h2 <-  ggplot(final.df, aes(x=maxcor_LFMMU_temp, y=LEA3.2_lfmm2_AUCPR_temp_neutSNPs, color=as.factor(N_traits))) + geom_point() + ggtheme + scale_color_viridis(option="mako", begin=.20, end=.80, discrete=TRUE) +  ylim(0,1) + ylab("AUC-PR temp") + geom_smooth() 

grid.arrange(g, h, i, nrow=2, ncol=2, layout_matrix=matrix(c(1,1,2,3), nrow=2, byrow=TRUE))
```

```{r}
j <- ggplot(final.df, aes(x=abs(cor_LFMMU1_temp), y=LEA3.2_lfmm2_TPR_temp, color=as.factor(N_traits))) + geom_point() + ggtheme + scale_color_viridis(option="mako", begin=.20, end=.80, discrete=TRUE) +  ylim(0,1) + ylab("TPR temp") + geom_smooth()  + xlim(0,1)

k <- ggplot(final.df, aes(x=abs(cor_LFMMU1_sal), y=LEA3.2_lfmm2_TPR_sal, color=as.factor(N_traits))) + geom_point() + ggtheme + scale_color_viridis(option="mako", begin=.20, end=.80, discrete=TRUE) + geom_smooth() + ylim(0,1)+ ylab("TPR Env2")+ xlim(0,1)

l <- ggplot(final.df, aes(x=abs(cor_LFMMU1_temp), y=LEA3.2_lfmm2_FDR_neutSNPs_temp, color=as.factor(N_traits))) + geom_point() + ggtheme + scale_color_viridis(option="mako", begin=.20, end=.80, discrete=TRUE) +  ylim(0,1) + ylab("FDR temp") + geom_smooth() + xlim(0,1)

  
m <- ggplot(final.df, aes(x=abs(cor_LFMMU1_sal), y=LEA3.2_lfmm2_FDR_neutSNPs_sal, color=as.factor(N_traits))) + geom_point() + ggtheme + scale_color_viridis(option="mako", begin=.20, end=.80, discrete=TRUE) + geom_smooth() + ylim(0,1)+ ylab("FDR Env2")+ xlim(0,1)


grid.arrange(j, k, l, m, nrow=2)

#ggplot(comparemethods.AUC.temp, aes(x=abs(cor_LFMMU2_temp), y=AUCPR, color=method, shape=method)) + geom_point() + ggtheme + scale_color_viridis(option="mako", discrete=TRUE) 

#ggplot(comparemethods.AUC.temp, aes(x=abs(cor_PC1_temp), y=AUCPR, color=method, shape=method)) + geom_point() + ggtheme + scale_color_viridis(option="mako", begin=.20, end=.80, discrete=TRUE) 


#ggplot(comparemethods.AUC.sal, aes(x=abs(cor_PC1_sal), y=AUCPR, color=method, shape=method)) + geom_point() + ggtheme + scale_color_viridis(option="mako", begin=.20, end=.80, discrete=TRUE) 

#ggplot(comparemethods.AUC.sal, aes(x=abs(cor_LFMMU2_sal), y=AUCPR, color=method, shape=method)) + geom_point() + ggtheme + scale_color_viridis(option="mako", discrete=TRUE) 

```

Is there a relationship between phenotypic clines and number of outliers?
```{r}
will be able to do with new outputs
```

## Can redundancy analysis identify mutations with pleiotropic effects?
```{r}
will be able to do with new outputs
```

## Under what conditions can redundancy analysis predict an individual’s optimal environment, without knowledge of the genetic architecture of adaptation?

```{r}
n <- ggplot(final.df, aes(x=cor_RDA20000_RDloadings_tempPhen, fill=demog_level)) + geom_histogram(color="grey30") + ggtheme + xlim(0,1) + xlab("Cor(temp. phenotype, RDA predicted temp. phenotype)")
o <- ggplot(final.df, aes(x=cor_RDA20000_RDloadings_salPhen, fill=demog_level)) + geom_histogram(color="grey30") + ggtheme + xlim(0,1) + xlab("Cor(Env2 phenotype, RDA predicted Env2 phenotype)")

grid.arrange(n, o, nrow=2)

#hist(final.df$cor_RDA20000_RDloadings_tempPhen, xlim=c(0,1), breaks=seq(0,1,0.05))
#hist(final.df$cor_RDA20000_RDloadings_salPhen, xlim=c(0,1), breaks=seq(0,1,0.05))
```

```{r}
n1 <- ggplot(final.df, aes(x=cor_RDA20000_RDloadings_tempPhen, fill=arch_level)) + geom_histogram(color="grey30") + ggtheme + xlim(0,1) + xlab("Cor(temp. phenotype, RDA predicted temp. phenotype)")
o1 <- ggplot(final.df, aes(x=cor_RDA20000_RDloadings_salPhen, fill=arch_level)) + geom_histogram(color="grey30") + ggtheme + xlim(0,1) + xlab("Cor(Env2 phenotype, RDA predicted Env2 phenotype)")

grid.arrange(n1, o1, nrow=2)

#hist(final.df$cor_RDA20000_RDloadings_tempPhen, xlim=c(0,1), breaks=seq(0,1,0.05))
#hist(final.df$cor_RDA20000_RDloadings_salPhen, xlim=c(0,1), breaks=seq(0,1,0.05))
```