---
title: "Non-clinal sims analysis"
author: "KE Lotterhos"
date: "9/20/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(ggplot2)
library(viridis)
library(gridExtra)
library(tidyr)
library(dplyr)
#library(tidyverse)
outputs <- "figures_20220428/"
```

`setwd("~/Documents/GitHub/MVP-NonClinalAF")`
setwd("/work/lotterhos/MVP-NonClinalAF")
## Read in data

```{r cars}
out.df <- read.table(file = "summary_20220428.txt", header=TRUE)
head(out.df)
tail(out.df)



load("src/0b-final_params-20220428.RData")
sims.df <- final
head(sims.df)

final.df <- merge(sims.df, out.df, all.x=TRUE)
head(final.df)
(vars <- t(final.df[1,]))

ref <- final.df[,c("seed", "level", "num_causal_postfilter")]

count <- data.frame(notNA=tapply(final.df$K, final.df$level, 
       function(x){sum(!is.na(x))}
       ))
count

sum(is.na(final.df$K))


levels(final.df$demog_level_sub)
levels(final.df$arch_level_sub)
#final.df %>% filter(demog_level=="Est-Clines" & 
#                      demog_level_sub=="N-equal_m-constant" & 
#                    arch_level_sub=="2-trait-pleiotropy-equal-S" &
#                      arch_level=="mod-polygenic")

#1231217
```

Copy graphs - no need to run this every time.
```{r, eval=FALSE, echo=FALSE}
#for (i in 1:225){ #unhash to run
  file <- paste0("cp sim_output_20220201/",final.df$seed[i],"*.pdf graphs_20220201/")
  system(file)
}

#for (i in 1:15){ #unhash to run
  file <- paste0("cp sim_output_20220201/",final.df$seed[i],"*1pop.pdf graphs2_20220201/")
  system(file)
}

#system("cp sim_output_20220201/1231217*.pdf graphs_1231217_estuary_pleiotropy_modpoly")
#system("cp sim_output_20220201/1231142*.pdf graphs_1231217_estuary_pleiotropy_modpoly")
#system("cp sim_output_20220201/1231442*.pdf graphs_1231217_estuary_pleiotropy_modpoly")
system("cp sim_output_20220201/1231667*.pdf graphs_1231217_estuary_pleiotropy_modpoly")

#1231892
#1232117
#1232342
```

Seeds needed for 10x2 demography graph are the first 15 seeds

GGtheme
```{r}

ggtheme <- theme_bw() + theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank(), panel.background = element_blank(), panel.border=element_blank(), axis.line = element_line(colour="grey30"), axis.title = element_text(colour="grey20"), axis.text = (element_text(colour="grey30")), legend.title = element_text(colour="grey20"), legend.text = element_text(colour="grey30"))

```

## Color schemes
`viridis` for SS, SS-Mtn, Estuary
`mako` for demography sub level (m and N)
`magma` for arch_level_sub


## order levels
```{r}
str(final.df$arch_level)
levels(final.df$arch_level)
final.df$arch_level <- factor(final.df$arch_level, levels=c( "oliogenic", "mod-polygenic", "highly-polygenic"), ordered=TRUE)

levels(final.df$arch_level) <- c("oligogenic", "mod.\npolygenic", "highly\npolygenic")
str(final.df$arch_level) 


levels(final.df$arch_level_sub)

final.df$arch_level_sub <-
  factor(final.df$arch_level_sub,
  levels=c("1-trait", "2-trait-no-pleiotropy-equal-S",
           "2-trait-pleiotropy-equal-S",
           "2-trait-no-pleiotropy-unequal-S",
           "2-trait-pleiotropy-unequal-S"), ordered=TRUE)

levels(final.df$arch_level_sub) <- c("1 trait", "2 traits, no pleiotropy, equal S",
           "2 traits, pleiotropy, equal S",
           "2 traits, no pleiotropy, unequal S",
           "2 traits, pleiotropy, unequal S")


levels(final.df$arch_level_sub) <- c("1 trait", "2 traits\nno pleiotropy\nequal S",
           "2 traits\npleiotropy\nequal S",
           "2 traits\nno pleiotropy\nunequal S",
           "2 traits\npleiotropy\nunequal S")

final.df$demog_level <- factor(final.df$demog_level, levels=c("SS-Clines", "SS-Mtn", "Est-Clines"), ordered=TRUE)

final.df$demog_level_sub<- factor(final.df$demog_level_sub, levels=c("N-equal_m-constant", "N-equal_m_breaks", "N-cline-N-to-S_m-constant", "N-cline-center-to-edge_m-constant",  "N-variable_m-variable"), ordered=TRUE)

final.df$demog_level_sub <- as.factor(final.df$demog_level_sub)

levels(final.df$demog_level_sub)

levels(final.df$demog_level_sub) <- c("N equal\nm constant", "N equal\nm breaks", "N latitude cline\nm constant", "N central cline\nm constant",  "N variable\nm variable") #warning, make sure in same order as previous code
levels(final.df$demog_level_sub)
```

## mean FST
```{r mean FST}
#hist(final.df$meanFst, xlim=c(0,1), breaks=seq(0,1,0.01), main="", xlab="Fst")

sum(is.na(final.df$meanFst))
min(final.df$meanFst)
max(final.df$meanFst)
sum(is.na(final.df$arch_level_sub))
sum(is.na(final.df$demog_level))

tapply(final.df$meanFst, list(final.df$demog_level, final.df$demog_level_sub), mean)

ggplot(final.df) + geom_histogram(aes(x=meanFst, fill=arch_level_sub), binwidth=0.01, color="grey") + ggtheme + xlim(0,0.5) + facet_grid(demog_level~.) + scale_fill_viridis(option="magma", discrete=TRUE, name="Pleiotropy level") + xlab("Fst") + ylab("Number of simulations")  + theme(legend.position="bottom") +  guides(fill = guide_legend(reverse = TRUE)) + ggtitle("A")

#I'm not sure where the missing values come from.

pdf(paste0(outputs,"FST_Demog.pdf"), width=8, height=6)
  ggplot(final.df) + geom_histogram(aes(x=meanFst, fill=demog_level_sub, position="stack"), binwidth=0.01, color="grey") + ggtheme + xlim(0,0.5) + facet_grid(demog_level~.) + scale_fill_viridis(option="mako", discrete=TRUE, name="Demography level") + xlab("Fst") + ylab("Number of simulations")  + theme(legend.position="bottom") 
  dev.off()

ggplot(final.df, aes(x=as.factor(demog_level_sub), y=meanFst, fill=demog_level)) + geom_boxplot(color="grey") + ggtheme + xlab("Number of traits") + ylab("overall FST") + scale_color_viridis(discrete=TRUE) + theme(legend.position = "none") + ylim(0,0.5) + ggtitle("C")

a<-tapply(final.df$meanFst,final.df$demog_name, mean, na.rm=TRUE)
b<- tapply(final.df$final_LA,final.df$demog_name, mean, na.rm=TRUE)
  
cbind(Fst=a,LA=b)

ggplot(final.df, aes(x=as.factor(N_traits), y=meanFst, fill=demog_level)) + geom_boxplot(color="grey") + ggtheme + xlab("Number of traits") + ylab("meanFst") + scale_color_viridis(discrete=TRUE) + theme(legend.position = "none") + ylim(0,1) + ggtitle("C")

```

## Degree local adaptation
```{r}
#hist(final.df$final_LA, xlim=c(0,1), breaks=seq(0,1,0.05), main="", xlab="Degree of local adaptation")

 #   ggplot(final.df) + geom_histogram(aes(x=final_LA, fill=as.factor(arch_level))) + ggtheme + xlim(0,1) + facet_grid(demog_level~.)
    
  #      ggplot(final.df) + geom_histogram(aes(x=final_LA, fill=as.factor(demog_level_sub))) + ggtheme + xlim(0,1) + facet_grid(demog_level~.)


      l1 <- ggplot(final.df) + geom_histogram(aes(x=final_LA, fill=arch_level_sub), binwidth=0.01, color="grey") + ggtheme + xlim(0,0.75) + facet_grid(demog_level~.) + scale_fill_viridis(option="magma", discrete=TRUE, name="Pleiotropy level") + xlab("Amount of local adaptation") + ylab("Number of simulations")  + theme(legend.position="bottom") +
  guides(fill = guide_legend(reverse = TRUE)) + ggtitle("A")
l1

ggplot(final.df) + geom_histogram(aes(x=final_LA, fill=demog_level_sub), binwidth=0.01, color="grey") + ggtheme + xlim(0,0.75) + facet_grid(demog_level~.) + scale_fill_viridis(option="mako", discrete=TRUE, name="Demography level") + xlab("Amount of local adaptation") + ylab("Number of simulations")  + theme(legend.position="bottom") +
  guides(fill = guide_legend(reverse = FALSE)) + ggtitle("A")

      l2<- ggplot(final.df) +
        geom_point(aes(x=subsamp_corr_phen_temp,
                       y=subsamp_corr_phen_sal,
                       color=final_LA,shape=demog_level)) + 
        ggtheme +
      scale_color_viridis(option="mako", name="Local\nAdaptation", begin=0.2, end=1) + xlim(0.25,1) + ylim(0.25,1) + xlab("Correlation(temp. phenotype, deme temperature)") + ylab("Correlation(Env2 phenotype, deme Env2)") + ggtitle("B")
  l2
  
  sum(is.na(final.df$subsamp_corr_phen_temp))
  sum(is.na(final.df$subsamp_corr_phen_sal))
      # in this case, the removed points have an NA because the "sal" trait was not adaptive if 450 simulations
  
pdf(paste0(outputs,"AmountLA.pdf"), width=8, height=8)
  grid.arrange(l1,l2, nrow=2)
dev.off()

```

## Number of loci
```{r}

numloc_df <- gather(final.df, key=filter, value=num_loci_causal, num_causal_prefilter, num_causal_postfilter)

numloc_df$filter <- factor(numloc_df$filter, levels = c("num_causal_prefilter", "num_causal_postfilter"), ordered=TRUE)

levels(numloc_df$filter) <- c("Pre MAF filter", "Post filter MAF > 0 .01")
str(numloc_df$filter)


tapply( numloc_df$num_loci_causal, list(numloc_df$filter,numloc_df$arch_level), mean, na.rm=TRUE)

str(numloc_df$arch_level)

m<- ggplot(numloc_df ) + geom_boxplot(aes(y=num_loci_causal,x = arch_level, fill=filter), color="grey70") + ggtheme + ylim(0,4000) + ylab("Number of causal loci") + xlab("Architecture") + scale_fill_viridis(option="mako", discrete=TRUE)
m

sum(is.na(final.df$num_neut_prefilter))
sum(is.na(final.df$num_neut_postfilter))

n <- ggplot(final.df) + geom_violin(aes(y=num_neut_postfilter, x=demog_level_sub, fill=demog_level), color="grey70") + ggtheme + ylab("Number of neutral loci\nafter sampling and MAF filter") + xlab("Demography") + guides(fill=guide_legend(title="Landscape"))
n

tapply(final.df$num_neut_postfilter, list(final.df$demog_level_sub), median, na.rm=TRUE)
tapply(final.df$num_neut_postfilter, list(final.df$demog_level_sub), mean, na.rm=TRUE)

pdf(paste0(outputs,"NumLociPrePostfilter.pdf"), width=7, height=8)
  grid.arrange(m, n, nrow=2)
dev.off()
```


## K (Number of populations)
```{r}
K_mod <- summary(aov(K~arch + demog_level_sub + demog_level + arch*demog_level_sub*demog_level, data=final.df))
K_mod
pervarK <- round(K_mod[[1]][,2]/sum(K_mod[[1]][,2]),2)
data.frame(name=rownames(K_mod[[1]]),pervarK)

pdf(paste0(outputs,"K_by_demog.pdf"), width=8, height=6)
ggplot(final.df) + geom_histogram(aes(x=K, fill=demog_level), binwidth=1, color="grey") + ggtheme + xlim(0,20) + facet_grid(demog_level_sub~.) + scale_fill_viridis(option="viridis", discrete=TRUE, name="Demography") + xlab("Number of population clusters (K)") + ylab("Number of simulations")  + theme(legend.position="bottom") +
  guides(fill = guide_legend(reverse = FALSE)) 
dev.off()
```

## Statistical models

#### MODELS FOR DIVERGENCE AND FOR AMOUNT LA
```{r}
levels(final.df$arch)
levels(final.df$demog_level)
levels(final.df$demog_level_sub)

LA_mod <- summary(aov(final_LA~arch + demog_level_sub + demog_level + arch*demog_level_sub*demog_level, data=final.df))
LA_mod
pervarLA <- round(LA_mod[[1]][,2]/sum(LA_mod[[1]][,2]),2)

FST_mod <- summary(aov(meanFst ~ arch + demog_level_sub + demog_level + arch*demog_level_sub*demog_level, data=final.df))
FST_mod
pervarFST <- round(FST_mod[[1]][,2]/sum(FST_mod[[1]][,2]),2)

data.frame(name=rownames(FST_mod[[1]]),pervarLA, pervarFST)
```


#### MODELS FOR AF CLINES
```{r}

stat_df <- gather(final.df, key=trait, value=cor_TPR, cor_TPR_temp, cor_TPR_sal)
tail(stat_df)

unique(stat_df$demog_level)
unique(stat_df$demog_level_sub)
unique(stat_df$demog_name)
unique(stat_df$arch_level_sub)
unique(stat_df$arch)

## A model with more hierarchy
# modelsum <- summary(aov(cor_TPR ~ trait + demog_level + demog_level_sub + arch_level_sub + arch_level + final_LA + trait*demog_level + trait*arch_level_sub, data=stat_df))
# 
# modelsum
# 
# pervar <- data.frame(variable = rownames(modelsum[[1]]), SS = modelsum[[1]][,2])
# 
# pervar$pervar <- pervar$SS/sum(pervar$SS)*100
# 
# pervar

## A model with less hierarchy
stat_df$trait <- factor(stat_df$trait)
stat_df$N_traits <- factor(stat_df$N_traits)
levels(stat_df$trait)
levels(stat_df$N_traits)
modelsum <- summary(aov(cor_TPR ~ trait + arch + demog + demog_level +  final_LA + N_traits , data=stat_df))
modelsum

pervar <- data.frame(variable = rownames(modelsum[[1]]), SS = modelsum[[1]][,2])

pervar$pervar <- round(pervar$SS/sum(pervar$SS)*100, 1)

pervar

final.df$N_traits <- factor(final.df$N_traits)
levels(final.df$N_traits)
modelsum <- summary(aov(cor_TPR_temp ~ arch + demog + demog_level + final_LA + arch*demog*demog_level, data=final.df))
pervar_temp <- data.frame(variable = rownames(modelsum[[1]]), SS = modelsum[[1]][,2])
pervar_temp$pervar <- round(pervar_temp$SS/sum(pervar_temp$SS)*100, 1)
pervar_temp

modelsum <- summary(aov(cor_TPR_sal ~ arch + demog + demog_level +  final_LA + arch*demog*demog_level, data=final.df))
pervar_sal <- data.frame(variable = rownames(modelsum[[1]]), SS = modelsum[[1]][,2])
pervar_sal$pervar <- round(pervar_sal$SS/sum(pervar_sal$SS)*100, 1)
pervar_sal

data.frame(variable = pervar_temp$variable, temp_AFcors = pervar_temp$pervar, sal_AFcors = pervar_sal$pervar)
```

# How does demography and genetic architecture contribute to the evolution of phenotypic clines without clines in causal allele frequencies?

## Number of loci with correlations - demog
```{r}
#boxplot(final.df$cor_TPR_temp~final.df$demog_level, ylim=c(0,1))




a<- ggplot(final.df, aes(x=as.factor(N_traits), y=cor_TPR_temp, fill=demog_level)) + geom_boxplot(color="grey") + ggtheme + xlab("Number of traits") + ylab("Proportion of causal temp. \n loci with significant clines") + scale_color_viridis(discrete=TRUE) + theme(legend.position = "none") + ylim(0,1) + ggtitle("C")


forb <- final.df[final.df$N_traits==2,]
forb$N_traits <- factor(forb$N_traits)
b<- ggplot(forb, aes(x=N_traits, y=cor_TPR_sal, fill=demog_level)) + geom_boxplot(color="grey") + ggtheme + xlab("Number of traits") + ylab("Proportion of causal Env2 \n loci with significant clines") + scale_color_viridis(discrete=TRUE) + ylim(0,1) + guides(fill=guide_legend(title="Landscape")) + ggtitle("D")



### Pleiotropy vs no pleiotropy

subdf_arch <- final.df[final.df$N_traits==2,]

unique(final.df$SIGMA_K_1)
unique(subdf_arch$SIGMA_K_2)
subdf_arch$SIGMA_K_2 <- as.factor(subdf_arch$SIGMA_K_2)

my_col_arch <- c(viridis(10)[3], viridis(10)[7])

f1 <-   ggplot(subdf_arch, aes(x=arch_level, y=cor_TPR_temp, fill=as.factor(ispleiotropy))) + geom_boxplot(color="grey") + ggtheme + xlab("") + ylab("Proportion of causal temp. \n loci with significant clines") + scale_fill_manual( values=my_col_arch, labels=c("No pleiotropy", "Pleiotropy"), name="2-trait architectures")  + ylim(0,1) #+ theme(legend.title=element_text("2-trait\narchitectures"))
f1

f2 <-   ggplot(subdf_arch, aes(x=arch_level, y=cor_TPR_sal, fill=as.factor(ispleiotropy))) + geom_boxplot(color="grey") + ggtheme + xlab("") + ylab("Proportion of causal sal. \n loci with significant clines") + scale_fill_manual( values=my_col_arch, labels=c("No pleiotropy", "Pleiotropy"), name="2-trait architectures")  + ylim(0,1) #+ theme(legend.title=element_text("2-trait\narchitectures"))
f2

# Combine data together for a single plot f1/f2
#unique(subdf_arch$SIGMA_K_2)

pleiotropy.df <- gather(subdf_arch, key=trait, value=cor_TPR, cor_TPR_temp, cor_TPR_sal)

fboth <-   ggplot(pleiotropy.df, aes(x=arch_level, y=cor_TPR, fill=as.factor(ispleiotropy))) + geom_boxplot(color="grey") + ggtheme + xlab("") + ylab("Proportion of causal \n loci with significant clines") + scale_fill_manual( values=my_col_arch, labels=c("No pleiotropy", "Pleiotropy"), name="2-trait architectures")  + ylim(0,1) + ggtitle("B")






## Phenotypic clines vs. af clines

### For each trait, I want to plot the (correlation between the trait and environment) vs. (correlation between af and environment)


f3 <- ggplot(final.df) + geom_point(aes(x=subsamp_corr_phen_temp, y=cor_TPR_temp, color=demog_level, shape=arch_level), alpha=0.6, fill="grey") +  geom_point(aes(x=subsamp_corr_phen_sal, y=cor_TPR_sal, color=demog_level, shape=arch_level), alpha=0.6, fill="grey") + ggtheme + ylim(0,1) + xlim(0,1)  +  xlab("Correlation(phenotype, deme environment)") + ylab("Proportion of causal \n loci with significant clines") + scale_color_viridis(name="Landscape", discrete=TRUE) + scale_shape_manual(values=c(0,23, 16),name="Genic level") + ggtitle("A")

pdf(paste0(outputs,"WhatDrivesNonClines.pdf"), width=6, height=8)
grid.arrange(f3, fboth,a,b, nrow=3, ncol=2, layout_matrix=matrix(c(1,1,2,2,3,4), nrow=3, byrow=TRUE))
dev.off()

```

## Percent of Va explained by clinal loci

Note that this is a vast overestimate, because many causal loci were filtered out. So it's worth thinking about how to present this.

```{r}

pleiotropy.df2 <- gather(subdf_arch, key=trait, value=cor_VA_prop, cor_VA_temp_prop, cor_VA_sal_prop)

fboth2 <-   ggplot(pleiotropy.df2, aes(x=arch_level, y=cor_VA_prop, fill=as.factor(ispleiotropy))) + geom_boxplot(color="grey") + ggtheme + xlab("") + ylab("Proportion of explainable VA \n explained by loci with significant clines") + scale_fill_manual( values=my_col_arch, labels=c("No pleiotropy", "Pleiotropy"), name="2-trait architectures")  + ylim(0,1) + ggtitle("A")

a1<- ggplot(final.df, aes(x=as.factor(N_traits), y=cor_VA_temp_prop, fill=demog_level)) + geom_boxplot(color="grey") + ggtheme + xlab("Number of traits") + ylab("Proportion of explainable VA in temp. \n explained by loci with significant clines") + scale_color_viridis(discrete=TRUE) + theme(legend.position = "none")+ ylim(0,1) + ggtitle("B")

b1<- ggplot(forb, aes(x=N_traits, y=cor_VA_sal_prop, fill=demog_level)) + geom_boxplot(color="grey") + ggtheme + xlab("Number of traits") + ylab("Proportion of explainable VA in Env2 \n loci  explained by loci with significant clines") + scale_color_viridis(discrete=TRUE) + ylim(0,1) + ggtitle("C")


pdf(paste0(outputs,"WhatDrivesNonClines_VA.pdf"), width=8, height=8)
  grid.arrange(fboth2, a1,b1, nrow=2, ncol=2, layout_matrix=matrix(c(1,1,2,3), nrow=2, byrow=TRUE))
dev.off()
```

## Demography sub level
```{r}

unique(final.df$demog)
unique(final.df$demog_level)
unique(final.df$demog_level_sub)



#g<- ggplot(final.df, aes(x=as.factor(demog_level_sub), y=cor_TPR_temp, fill=as.factor(demog_level_sub))) + geom_boxplot(color="grey") + ggtheme + xlab("") + ylab("Proportion of causal temp. \n loci with significant clines") + scale_fill_viridis(option="magma",discrete=TRUE) + theme(legend.position = "none") + theme( axis.text.x=element_blank())
#g

#h <- ggplot(final.df, aes(x=as.factor(demog_level_sub), y=cor_TPR_sal, fill=demog_level)) + geom_boxplot(color="grey") + ggtheme + xlab("Demography sublevel") + ylab("Proportion of causal Env2 \n loci with significant clines") + scale_color_viridis(option="magma",discrete=TRUE) + theme(legend.position = "none") + theme( axis.text.x=element_text(angle=90,vjust=1, size = 10))
#grid.arrange(g,h, nrow=7, ncol=1, layout_matrix=matrix(c(1,1,2,2,2,2,2), nrow=7))



pleiotropy.df$demog_level_sub <- as.factor(pleiotropy.df$demog_level_sub)
levels(pleiotropy.df$demog_level_sub) <- c("N equal\nm constant", "N equal\nm breaks", "N latitude cline\nm constant", "N central cline\nm constant",  "N variable\nm variable") #warning, make sure in correct order




## Sub demograph - magma

d1 <- ggplot(pleiotropy.df, aes(x=as.factor(demog_level_sub), y=cor_TPR, fill=as.factor(demog_level_sub))) + geom_boxplot(color="grey") + ggtheme + xlab("Demography level") + ylab("Proportion of causal loci\nwith significant clines") + scale_fill_viridis(option="mako",discrete=TRUE) + theme(legend.position = "none") + theme( axis.text.x=element_text(angle=90,vjust=0.5, hjust=1, size = 10)) + ggtitle("A")
d1

## Strength of selection - mako
my_col_S <- c(magma(10)[5], magma(10)[8])

s1<- ggplot(subdf_arch, aes(x=arch_level, y=cor_TPR_temp, fill=as.factor(SIGMA_K_2))) + geom_boxplot(color="grey") + ggtheme + xlab("") + ylab("Proportion of causal temp. \n loci with significant clines") + scale_fill_manual( values=my_col_S, labels=c(
  expression(paste("Strong selection ", sigma[K]," = 0.5")), 
  expression(paste("Weak selection ", sigma[K]," = 4.0"))
  ), name="2-trait architectures")  + ylim(0,1) + xlab("Genetic architecture") + ggtitle("B")
s1

pdf(paste0(outputs,"WhatDrivesNonClines-subdemog-selection.pdf"), width=6, height=8)
  grid.arrange(d1, s1, nrow=2)
dev.off()
```



Changing strength of selection on trait 1 doesn't effect clines in trait 2
```{r}
 #+ theme(legend.title=element_text("2-trait\narchitectures"))

ggplot(subdf_arch, aes(x=arch_level, y=cor_TPR_sal, fill=as.factor(SIGMA_K_2))) + geom_boxplot(color="grey") + ggtheme + xlab("") + ylab("Proportion of causal sal. \n loci with significant clines") + scale_fill_manual( values=my_col_S, labels=c("Strong selection on other trait (sigma_K=0.5)", "Weak selection on other trait (sigma_K=4)"), name="2-trait architectures")  + ylim(0,1) #+ theme(legend.title=element_text("2-trait\narchitectures"))

```


## Distribution between structure and deme environment

```{r}

s1 <- ggplot(final.df) + geom_density(aes(x=abs(cor_PC1_temp), y=..scaled..), fill="cornflowerblue") + ggtheme + ggtitle("A") + xlab("abs(Correlation(PC1, temperature))")  + ylab(NULL)
s1

s2 <-ggplot(final.df) + geom_density(aes(x=abs(cor_PC2_temp), y=..scaled..), fill="cornflowerblue") + ggtheme + ggtitle("B") + xlab("abs(Correlation(PC2, temperature))") + ylab(NULL)
s2

s3 <-ggplot(final.df) + geom_density(aes(x=abs(cor_PC1_sal), y=..scaled..), fill="chartreuse3") + ggtheme + ggtitle("C") + xlab("abs(Correlation(PC1, Env2))") + ylab(NULL)
s3

s4 <-ggplot(final.df) + geom_density(aes(x=abs(cor_PC2_sal), y=..scaled..), fill="chartreuse3") + ggtheme + ggtitle("D") + xlab("abs(Correlation(PC2, Env2))") + ylab(NULL)
s4

sum(is.na(final.df$cor_PC1_temp))

sc1<- ggplot(final.df) + geom_histogram(aes(x=abs(cor_PC1_temp), fill=demog_level_sub, position="stack", ), binwidth=0.01, color="grey") + ggtheme + xlim(0,1) + facet_grid(demog_level~., scales = "free_y") + scale_fill_viridis(option="mako", discrete=TRUE, name="Demography level") + xlab("|Cor(PC1, deme temp)|") + ylab("Number of simulations")  + theme(legend.position="none") + ggtitle("A")

sc2<-ggplot(final.df) + geom_histogram(aes(x=abs(cor_PC2_temp), fill=demog_level_sub, position="stack"), binwidth=0.01, color="grey") + ggtheme + xlim(0,1) + facet_grid(demog_level~., scales = "free_y") + scale_fill_viridis(option="mako", discrete=TRUE, name="Demography level") + xlab("|Cor(PC2, deme temp)|") + ylab("Number of simulations")  + theme(legend.position="none") + ggtitle("B")

sc3<-ggplot(final.df) + geom_histogram(aes(x=abs(cor_PC1_sal), fill=demog_level_sub, position="stack", ), binwidth=0.01, color="grey") + ggtheme + xlim(0,1) + facet_grid(demog_level~., scales = "free_y") + scale_fill_viridis(option="mako", discrete=TRUE, name="Demography level") + xlab("|Cor(PC1, deme sal)|") + ylab("Number of simulations")  + theme(legend.position="none") + ggtitle("C")

# bw = "nrd", kernel = "rectangular", n=2^10, adjust=1/5,
sc4<-ggplot(final.df) + geom_histogram(aes(x=abs(cor_PC2_sal), fill=demog_level_sub, position="stack"), binwidth=0.01,  color="grey") + ggtheme + xlim(0,1) + facet_grid(demog_level~., scales = "free_y") + scale_fill_viridis(option="mako", discrete=TRUE, name="Demography level") + xlab("|Cor(PC2, deme sal)|") + ylab("Number of simulations")  + theme(legend.position="none") + ggtitle("D")

pdf(paste0(outputs,"StructureEnvCor.pdf"), width=6, height=6)
  grid.arrange(s1,s2, s3,s4, nrow=2)
dev.off()

pdf(paste0(outputs,"StructureEnvCor_byDemog.pdf"), width=13, height=6)
grid.arrange(s1 + facet_grid(~demog_level),
             s2 + facet_grid(~demog_level),
             s3 + facet_grid(~demog_level),
             s4 + facet_grid(~demog_level)
             )
dev.off()

pdf(paste0(outputs,"StructureEnvCor_facetDemog.pdf"), width=12, height=16)
    grid.arrange(sc1,sc2, sc3,sc4, nrow=2)
dev.off()

```

## Under what conditions can we reliably identify the alleles with effects on adaptive phenotypes using genetic-environment associations (latent factor mixed models, uncorrected correlations, and redundancy analysis)? 


## Compare methods AUC
```{r}

# Missing data for temp - 2 sims
sum(is.na(final.df$cor_AUCPR_temp_neutSNPs))
sum(is.na(final.df$LEA3.2_lfmm2_AUCPR_temp_neutSNPs))
sum(is.na(final.df$RDA_AUCPR_neutSNPs))

# Missing data for sal - 1 case (note missing data is expected when N_traits==1, because salinity not simulated)
sum(is.na(final.df$cor_AUCPR_sal_neutSNPs & final.df$N_traits==2))
sum(is.na(final.df$LEA3.2_lfmm2_AUCPR_sal_neutSNPs& final.df$N_traits==2))


comparemethods.AUC.temp <- gather(final.df, key=method, value=AUCPR, cor_AUCPR_temp_neutSNPs, LEA3.2_lfmm2_AUCPR_temp_neutSNPs, RDA_AUCPR_neutSNPs)

comparemethods.AUC.sal <- gather(final.df, key=method, value=AUCPR, cor_AUCPR_sal_neutSNPs, LEA3.2_lfmm2_AUCPR_sal_neutSNPs, RDA_AUCPR_neutSNPs)

whichcols <- c("method", "AUCPR", "cor_PC1_temp", "cor_PC1_sal","cor_LFMMU1_temp", "cor_LFMMU1_sal", "cor_LFMMU2_temp", "cor_LFMMU2_sal", "demog_level", "demog_level_sub", "arch_level", "arch_level_sub", "K")

comparemethods.AUC <- rbind(comparemethods.AUC.temp[,whichcols], comparemethods.AUC.sal[whichcols])


methodtype <- unlist(lapply(comparemethods.AUC$method, function(x){strsplit(x, "_")[[1]][1]}))
str(methodtype)
comparemethods.AUC$methodtype <- methodtype

# Get lower and upper bounds of box in boxplot
#tapply(comparemethods.AUC$AUCPR, as.factor(comparemethods.AUC$methodtype), quantile,p=0.25, na.rm=TRUE)
#tapply(comparemethods.AUC$AUCPR, as.factor(comparemethods.AUC$methodtype), quantile,p=0.75, na.rm=TRUE)

# By landscape
g <- ggplot(comparemethods.AUC, aes(x=as.factor(methodtype), y=AUCPR, fill=demog_level)) + geom_boxplot(color="grey") + ggtheme + xlab("method") + ylab("AUC-PR")   + theme( axis.text.x=element_text(angle=90,vjust=0.5, hjust=1, size = 10)) + scale_color_viridis(name="Landscape", discrete=TRUE) + guides(fill=guide_legend(title="Landscape"))

pdf(paste0(outputs,"AUC-PR-byLandscape.pdf"), width=5, height=4)
g
dev.off()

# as a violin plot (the boxplot is better)
#ggplot(comparemethods.AUC, aes(x=as.factor(methodtype), y=AUCPR, fill=demog_level)) + geom_violin(color="grey") + ggtheme + xlab("method") + ylab("AUC-PR")   + theme( axis.text.x=element_text(angle=90,vjust=0.5, hjust=1, size = 10))


# By demography - not a huge difference among methods
#ggplot(comparemethods.AUC, aes(x=as.factor(methodtype), y=AUCPR, fill=demog_level_sub)) + geom_boxplot(color="grey") + ggtheme + xlab("method") + ylab("AUC-PR")   + theme( axis.text.x=element_text(angle=90,vjust=0.5, hjust=1, size = 10))
```



## Compare methods FDR
```{r}

comparemethods.FDR.temp <- gather(final.df, key=method, value=FDR, cor_FDR_neutSNPs_temp, LEA3.2_lfmm2_FDR_neutSNPs_temp, RDA_FDR_neutSNPs)

comparemethods.FDR.sal <- gather(final.df, key=method, value=FDR, cor_FDR_neutSNPs_sal, LEA3.2_lfmm2_FDR_neutSNPs_sal, RDA_FDR_neutSNPs)

whichcols <- c( "seed", "level", "method", "FDR", "cor_PC1_temp", "cor_PC1_sal","cor_LFMMU1_temp", "cor_LFMMU1_sal", "cor_LFMMU2_temp", "cor_LFMMU2_sal")

comparemethods.FDR <- rbind(comparemethods.FDR.temp[,whichcols], comparemethods.FDR.sal[whichcols])
head(comparemethods.FDR[order(comparemethods.FDR$seed),])

methodtype <- unlist(lapply(comparemethods.FDR$method, function(x){strsplit(x, "_")[[1]][1]}))
str(methodtype)
comparemethods.FDR$methodtype <- methodtype
comparemethods.FDR <- comparemethods.FDR[order(comparemethods.FDR$seed),]
tail(comparemethods.FDR)


g2<- ggplot(comparemethods.FDR, aes(x=as.factor(methodtype), y=FDR, fill=methodtype)) + geom_violin(color="grey") + ggtheme + xlab("method") + ylab("FDR") + scale_fill_viridis(option="turbo", discrete=TRUE) + theme(legend.position = "none") + theme( axis.text.x=element_text(angle=90,vjust=0.5, hjust=1, size = 10))

Fdr1 <- ggplot(comparemethods.FDR[grep("LEA",comparemethods.FDR$method),], aes(x=FDR)) + geom_density(color="grey", fill="blue") + ggtheme + xlab("method") + xlab("FDR") + scale_fill_viridis(option="turbo", discrete=TRUE) + theme(legend.position = "none") + ggtitle("A) LFMM")

Fdr2 <- ggplot(comparemethods.FDR[grep("cor",comparemethods.FDR$method),], aes(x=FDR)) + geom_density(color="grey", fill="blue") + ggtheme + xlab("method") + xlab("FDR") + scale_fill_viridis(option="turbo", discrete=TRUE) + theme(legend.position = "none") + ggtitle("B) Correlation")

Fdr3 <- ggplot(comparemethods.FDR[grep("RDA",comparemethods.FDR$method),], aes(x=FDR)) + geom_density(color="grey", fill="blue") + ggtheme + xlab("method") + xlab("FDR") + scale_fill_viridis(option="turbo", discrete=TRUE) + theme(legend.position = "none") + ggtitle("C) RDA")

pdf(paste0(outputs,"FDRdensityPerMethod.pdf"), width=5, height=7)

grid.arrange(Fdr1, Fdr2, Fdr3, nrow=3)

dev.off()

```

# Trying to understand demography effect on lfmm FDR
```{r}

l1 <- ggplot(final.df) + geom_boxplot(aes(y=LEA3.2_lfmm2_FDR_neutSNPs_temp, x=demog_level_sub, fill=demog_level), color="grey") + ggtheme + guides(fill=guide_legend(title="Landscape")) + xlab("Demography") + ylab("FDR for LFMM (temperature model)")  + theme(legend.position="none") + ylim(0,1)

l2 <- ggplot(final.df) + geom_boxplot(aes(y=LEA3.2_lfmm2_FDR_neutSNPs_sal, x=demog_level_sub, fill=demog_level), color="grey") + ggtheme + guides(fill=guide_legend(title="Landscape")) + xlab("Demography") + ylab("FDR for LFMM (Env2 model)")  + theme(legend.position="none") + ylim(0,1)


#ggplot(final.df) + geom_violin(aes(y=LEA3.2_lfmm2_AUCPR_temp_neutSNPs, x=demog_level_sub, fill=demog_level)) + ggtheme + guides(fill=guide_legend(title="Landscape")) + xlab("Demography") + ylab("")

l3 <- ggplot(final.df) + geom_boxplot(aes(y=abs(cor_LFMMU1_temp), x=demog_level_sub, fill=demog_level), color="grey") + ggtheme + guides(fill=guide_legend(title="Landscape")) + xlab("Demography") + ylab("Correlation (LFMM U1, deme temperature)")  + theme(legend.position="none") + ylim(0,1)


l4 <- ggplot(final.df) + geom_boxplot(aes(y=abs(cor_LFMMU2_temp), x=demog_level_sub, fill=demog_level), color="grey") + ggtheme  + xlab("Demography") + ylab("Correlation (LFMM U2, deme temperature)")   + theme(legend.position="none") + theme(legend.position="bottom") + ylim(0,1) + guides(fill=guide_legend(title="Landscape"))


l5 <- ggplot(final.df) + geom_boxplot(aes(y=abs(cor_LFMMU2_sal), x=demog_level_sub, fill=demog_level), color="grey") + ggtheme + guides(fill=guide_legend(title="Landscape")) + xlab("Demography") + ylab("Correlation (LFMM U1, deme Env2)")   + theme(legend.position="none") + ylim(0,1)

l6 <- ggplot(final.df) + geom_boxplot(aes(y=abs(cor_LFMMU2_sal), x=demog_level_sub, fill=demog_level), color="grey") + ggtheme + guides(fill=guide_legend(title="Landscape")) + xlab("Demography") + ylab("Correlation (LFMM U2, deme Env2)") + theme(legend.position="bottom") + ylim(0,1)

pdf(paste0(outputs,"FDR_LFMM.pdf"), width=10, height=12)

grid.arrange(l1,l2, l3, l5, l4,l6, ncol=2)

dev.off()

```

Old code for trying to understand demograpy effect in lfmm
```
ggplot(final.df) + geom_boxplot(aes(y=abs(cor_LFMMU2_temp), x=demog_level_sub)) + ggtheme

ggplot(final.df) + geom_boxplot(aes(y=abs(cor_LFMMU1_sal), x=demog_level_sub)) + ggtheme 
ggplot(final.df) + geom_boxplot(aes(y=abs(cor_LFMMU2_sal), x=demog_level_sub)) + ggtheme



ggplot(final.df) + geom_boxplot(aes(y=abs(cor_PC1_sal), x=demog_level_sub)) + ggtheme

fordemog <- gather(final.df, key=Uaxis, value=cor_LFMM_temp, cor_LFMMU1_temp, cor_LFMMU2_temp)

a1 <- ggplot(final.df) + geom_boxplot(aes(y=abs(cor_PC1_temp), x=demog_level_sub)) + ggtheme + ylim(0,1)
a2 <- ggplot(fordemog) + geom_boxplot(aes(y=abs(cor_LFMM_temp), x=demog_level_sub, fill=Uaxis)) + ggtheme +  theme(legend.position="bottom") + ylab("Cor(Latent factor, temp)") + ylim(0,1)
a3 <- ggplot(final.df) + geom_boxplot(aes(y=abs(LEA3.2_lfmm2_FPR_neutSNPs_temp), x=demog_level_sub)) + ggtheme + ylab("FPR") + ylim(0,1)
a4 <- ggplot(final.df) + geom_boxplot(aes(y=abs(LEA3.2_lfmm2_TPR_temp), x=demog_level_sub)) + ggtheme + ylab("Power") + ylim(0,1)
a5 <- ggplot(final.df) + geom_violin(aes(y=abs(LEA3.2_lfmm2_FDR_neutSNPs_temp), x=demog_level_sub)) + ggtheme + ylab("FDR") + ylim(0,1)
a6 <- ggplot(final.df) + geom_violin(aes(y=abs(LEA3.2_lfmm2_AUCPR_temp_neutSNPs), x=demog_level_sub)) + ggtheme + ylab("FDR") + ylim(0,1)

grid.arrange(a1,a2,a3, a4 ,a5, nrow=5, heights=c(1,1.5,1,1,1))

fordemog <- gather(final.df, key=Uaxis, value=cor_LFMM_sal, cor_LFMMU1_sal, cor_LFMMU2_sal)

e1 <- ggplot(final.df) + geom_boxplot(aes(y=abs(cor_PC1_sal), x=demog_level_sub)) + ggtheme + ylim(0,1)
e2 <- ggplot(fordemog) + geom_boxplot(aes(y=abs(cor_LFMM_sal), x=demog_level_sub, fill=Uaxis)) + ggtheme +  theme(legend.position="bottom") + ylab("Cor(Latent factor, sal)") + ylim(0,1)
e3 <- ggplot(final.df) + geom_boxplot(aes(y=abs(LEA3.2_lfmm2_FPR_neutSNPs_sal), x=demog_level_sub)) + ggtheme + ylab("FPR") + ylim(0,1)
e4 <- ggplot(final.df) + geom_boxplot(aes(y=abs(LEA3.2_lfmm2_TPR_sal), x=demog_level_sub)) + ggtheme + ylab("Power") + ylim(0,1)
e5 <- ggplot(final.df) + geom_violin(aes(y=abs(LEA3.2_lfmm2_FDR_neutSNPs_sal), x=demog_level_sub)) + ggtheme + ylab("FDR") + ylim(0,1)
e6 <-  ggplot(final.df) + geom_violin(aes(y=abs(LEA3.2_lfmm2_AUCPR_sal_neutSNPs), x=demog_level_sub)) + ggtheme + ylab("FDR") + ylim(0,1)

grid.arrange(a1,e1,a2,e2,a3,e3,a4, e4 ,a5,e5,a6,e6, nrow=6, ncol=2, heights=c(1,1.5,1,1,1,1))

grid.arrange(a1,e1,a2,e2, a6,e6, nrow=3, ncol=2, heights=c(1,1.5,1))


ggplot(comparemethods.FDR
CHECK FDR OUTPUTUS for m breaks- NOT MAKING SENSE



#ggplot(final.df) + geom_violin(aes(y=LEA3.2_lfmm2_AUCPR_temp_neutSNPs, x=demog_level_sub, fill=demog_level)) + ggtheme + ylab("AUC PR") 

#ggplot(final.df) + geom_violin(aes(y=LEA3.2_lfmm2_AUCPR_sal_neutSNPs, x=demog_level_sub, fill=demog_level)) + ggtheme + ylab("AUC PR")
```

# LFMM - AUCPR vs STRUCTURE
```{r}

#ggplot(final.df) + geom_point(aes(x=cor_TPR_temp,y=LEA3.2_lfmm2_AUCPR_temp_neutSNPs, color=demog_level_sub, shape=demog_level_sub)) + ggtheme + xlab("Proportion of causal temp loci with clines") + ylab("AUC-PR")

#ggplot(final.df) + geom_point(aes(x=abs(cor_PC1_temp),y=LEA3.2_lfmm2_AUCPR_temp_neutSNPs, color=demog_level_sub, shape=demog_level_sub)) + ggtheme + xlab("cor(pc1,temp)") + ylab("AUC-PR")

#ggplot(final.df) + geom_point(aes(x=cor_TPR_sal,y=LEA3.2_lfmm2_AUCPR_sal_neutSNPs, color=demog_level_sub, shape=demog_level_sub)) + ggtheme + xlab("Proportion of causal sal loci with clines") + ylab("AUC-PR")

#ggplot(final.df) + geom_point(aes(x=abs(cor_PC1_sal),y=LEA3.2_lfmm2_AUCPR_sal_neutSNPs, color=demog_level_sub, shape=demog_level_sub)) + ggtheme + xlab("cor(pc1,sal)") + ylab("AUC-PR")


k1 <- ggplot(final.df) + geom_point(aes(x=abs(cor_LFMMU1_temp),y=LEA3.2_lfmm2_AUCPR_temp_neutSNPs, color=demog_level_sub, shape=demog_level_sub)) + ggtheme + xlab("cor(LFMM U1, Deme Temperature)") + ylab("AUC-PR") + geom_smooth(aes(x=abs(cor_LFMMU1_temp),y=LEA3.2_lfmm2_AUCPR_temp_neutSNPs, color=demog_level_sub, fill=demog_level_sub), alpha=0.2, method="loess", span=1) + ylim(0,1)  +xlim(0,1)

k2 <- ggplot(final.df) + geom_point(aes(x=abs(cor_LFMMU1_sal),y=LEA3.2_lfmm2_AUCPR_sal_neutSNPs, color=demog_level_sub, shape=demog_level_sub)) + ggtheme + xlab("cor(LFMM U1, Deme Env2)") + ylab("AUC-PR") + geom_smooth(aes(x=abs(cor_LFMMU1_sal),y=LEA3.2_lfmm2_AUCPR_sal_neutSNPs, color=demog_level_sub, fill=demog_level_sub), alpha=0.2, method="loess", span=1) + ylim(0,1) +xlim(0,1)

pdf(paste0(outputs,"LFMM-AUCPRvsStructure.pdf"), width=5, height=8)

grid.arrange(k1,k2, nrow=2)

dev.off()
```

```{r}
ISSUES
noutliers <- final.df$LEA3.2_lfmm2_mlog10P_tempenv_noutliers
numTP <- final.df$num_causal_postfilter*final.df$LEA3.2_lfmm2_TPR_temp
str(numTP)
numFP <- final.df$num_neut*final.df$LEA3.2_lfmm2_FPR_neutSNPs_temp
plot(noutliers, numTP + numFP)
abline(0,1)
WHAT I CALCULATED IS NOT THE FDR!

  head(final.df)
  
  NOT GOING TO LINE UP B/C NEUT INCLUDES LINKED LOCI
fdr <- numFP/(numTP+numFP)
plot(fdr, final.df$LEA3.2_lfmm2_FDR_neutSNPs_temp)
abline(0,1)
hist(fdr)
#tapply(final.df$LEA3.2_lfmm2_TPR_temp, final.df$demog_level_sub, function(x){sum(!is.na(x))})
```


LFMM has a bimodal distribution of FDR. At low FDR, pop structure is appropriately accounted for. But at high FDR, is pop structure undercorrected or overcorrected?

* If undercorrected, then expect high FDR when the correlation between the latent factor and temp env is very low. 

* If overcorrected, then expect high FDR when the correlation between the latent factor and temp env is very high.



### Stat models LFMM

I don't partciularly trust these models, as some relationships are not linear and likely violate the assumption of the models.
```{r}

# Temperature models
summary(lm(final.df$LEA3.2_lfmm2_FDR_neutSNPs_temp~
             abs(final.df$cor_LFMMU1_temp)+
             abs(final.df$cor_LFMMU2_temp)+
             abs(cor_PC1_temp)
             , data=final.df))
# FDR increases when the correation between structure and environment increases, but it's bimodal as shown in the graph below

summary(lm(final.df$LEA3.2_lfmm2_TPR_temp~
             abs(final.df$cor_LFMMU1_temp)+
             abs(final.df$cor_LFMMU2_temp)+
             abs(cor_PC1_temp)
             , data=final.df))
# TPR decreases when the correlation between structure and environment increases

summary(lm(final.df$LEA3.2_lfmm2_FPR_neutSNPs_temp~
             abs(final.df$cor_LFMMU1_temp)+
             abs(final.df$cor_LFMMU2_temp)+
             abs(cor_PC1_temp)
             , data=final.df))
# FPR increases when the correlation between PC structure and environment increases


# Salinity models
summary(lm(final.df$LEA3.2_lfmm2_FDR_neutSNPs_sal~
             abs(final.df$cor_LFMMU1_sal)+
             abs(final.df$cor_LFMMU2_sal)+
             abs(cor_PC1_sal)
             , data=final.df))
# FDR decreases when the correation between structure and environment increases

summary(lm(final.df$LEA3.2_lfmm2_TPR_sal~
             abs(final.df$cor_LFMMU1_sal)+
             abs(final.df$cor_LFMMU2_sal)+
             abs(cor_PC1_sal)
             , data=final.df))
# TPR not related to structure


summary(lm(final.df$LEA3.2_lfmm2_FPR_neutSNPs_sal~
             abs(final.df$cor_LFMMU1_sal)+
             abs(final.df$cor_LFMMU2_sal)+
             abs(cor_PC1_sal)
             , data=final.df))
# FPR decreases when the correation between structure and environment increases


j1 <- ggplot(final.df[order(final.df$K),]) + 
  geom_point(aes(x=abs(cor_LFMMU1_temp), 
                 y=LEA3.2_lfmm2_FDR_neutSNPs_temp, 
                 color=abs(cor_LFMMU2_temp))) + 
  geom_point(aes(x=abs(cor_LFMMU1_sal), 
                 y=LEA3.2_lfmm2_FDR_neutSNPs_sal, 
                 color=abs(cor_LFMMU2_sal))) + ggtheme +
 scale_color_viridis(option="mako", begin=.20, end=.80, name="Correlation\n(latent factor 2,\nenvironment)") +
  ylab("False discovery rate") + xlab("Correlation\n(latent factor 1, environment)")
j1

```

## Is there a relationship between phenotypic clines and number of outliers?

Does the proportion of loci with clines to the environment increase when the phenotypic clines increase?

This figure isn't really needed, the one for the main paper is better
```{r}

c11 <- ggplot(final.df) + geom_point(aes(x=subsamp_corr_phen_temp,
                                  y=cor_TPR_temp,
                                  color=demog_level)) +
  xlim(0,1) + ylim(0,1) + ggtheme + xlab("Cor(temp. phenotype, deme temp.)") + ylab("Proportion of causal temp.\nloci with significant clines")

c12 <- ggplot(final.df) + geom_point(aes(x=subsamp_corr_phen_sal,
                                  y=cor_TPR_sal,
                                  color=demog_level)) +
  xlim(0,1) + ylim(0,1) + ggtheme + xlab("Cor(Env2 phenotype, deme Env2)") + ylab("Proportion of causal Env2\nloci with significant clines")

#pdf(paste0(outputs,"PhenCorVsPropClines.pdf"), width=8, height=8)
  grid.arrange(c11, c12)
#dev.off()

```

## Can redundancy analysis identify mutations with pleiotropic effects?
```{r}
#will be able to do with new outputs

par(mfrow=c(2,2))
hist(final.df$RDA_RDAmutpred_cor_tempEffect)
hist(final.df$RDA_RDAmutpred_cor_salEffect)

hist(final.df$cor_RDA20000_RDloadings_tempPhen)
hist(final.df$cor_RDA20000_RDloadings_salPhen)


b1<- final.df$seed[which(final.df$RDA_RDAmutpred_cor_tempEffect < 0)]
b2<- final.df$seed[which(final.df$RDA_RDAmutpred_cor_salEffect < 0)]
b3<- final.df$seed[which(final.df$cor_RDA20000_RDloadings_tempPhen < 0)]

failedsims <- unique(c(b1,b2,b3))
failedsims
system("mkdir graphs_badRDA20220201/")

for (i in failedsims){
  file <- paste0("cp sim_output_20220201/",i,"*.pdf graphs_badRDA20220201/")
  system(file)
}

length(failedsims)
# method failed for 20 sims

# Characteristics of failed sims
final.df$arch[which(final.df$seed %in% failedsims)]
final.df$demog_level[which(final.df$seed %in% failedsims)]
final.df$demog_level_sub[which(final.df$seed %in% failedsims)]
final.df$N_traits[which(final.df$seed %in% failedsims)]
final.df$K[which(final.df$seed %in% failedsims)]

# Not sure why it didn't work for these sims
```

## Under what conditions can redundancy analysis predict an individual’s optimal environment, without knowledge of the genetic architecture of adaptation?

```{r}
n <- ggplot(final.df, aes(x=cor_RDA20000_RDloadings_tempPhen, fill=demog_level)) + geom_histogram(color="grey40") + ggtheme + xlim(-1,1) + xlab("Cor(temp. phenotype,\nRDA-predicted temp. phenotype)") + guides(fill=guide_legend(title="Landscape")) + ggtitle("A) Temperature phenotype") + theme(legend.position = "none")
n

o <- ggplot(final.df, aes(x=cor_RDA20000_RDloadings_salPhen, fill=demog_level)) + geom_histogram(color="grey40") + ggtheme + xlim(-1,1) + xlab("Cor(Env2 phenotype,\nRDA-predicted Env2 phenotype)") + guides(fill=guide_legend(title="Landscape")) + ggtitle("B) Env. 2 phenotype") + theme(legend.position="bottom")
o


g1<- ggplot(final.df, aes(x=RDA_RDAmutpred_cor_tempEffect, fill=arch_level)) + geom_histogram(color="grey40") + ggtheme + xlim(-1,1) + xlab("Cor(mutation effect on temp,\nRDA-predicted temp. effect)") + guides(fill=guide_legend(title="Genetic\nArchitecture")) + scale_fill_viridis(option="mako", discrete=TRUE) + theme(legend.position = "none") + ggtitle("C) Temperature mutation effect")
g1 

g2 <- ggplot(final.df, aes(x=RDA_RDAmutpred_cor_salEffect, fill=arch_level)) + geom_histogram(color="grey40") + ggtheme + xlim(-1,1) + xlab("Cor(mutation effect on Env2,\nRDA-predicted Env2 effect)") + guides(fill=guide_legend(title="Genetic\nArchitecture")) + scale_fill_viridis(option="mako", discrete=TRUE) + theme(legend.position="bottom") + ggtitle("D) Env2 mutation effect")
g2

pdf(paste0(outputs,"RDA-Landscape-hist.pdf"), width=9, height=7)
  grid.arrange(arrangeGrob(n, o,  nrow=2),
               arrangeGrob(g1, g2, nrow=2),
               widths=c(5,5))
dev.off()


#hist(final.df$cor_RDA20000_RDloadings_tempPhen, xlim=c(0,1), breaks=seq(0,1,0.05))
#hist(final.df$cor_RDA20000_RDloadings_salPhen, xlim=c(0,1), breaks=seq(0,1,0.05))
```

Do other factors affect the correlation between mutation loading and mutation effect on phenotype?
```{r}

ggplot(final.df, aes(x=RDA_RDAmutpred_cor_tempEffect, fill=demog_level)) + geom_histogram(color="grey40") + ggtheme + xlim(0,1) + xlab("Cor(mutation effect on temp,\nRDA-predicted temp. effect)") + guides(fill=guide_legend(title="Landscape")) 

ggplot(final.df, aes(x=RDA_RDAmutpred_cor_tempEffect, fill=demog_level_sub)) + geom_histogram(color="grey40") + ggtheme + xlim(-1,1) + xlab("Cor(mutation effect on temp,\nRDA-predicted temp. effect)") + guides(fill=guide_legend(title="Landscape"))

ggplot(final.df, aes(x=RDA_RDAmutpred_cor_tempEffect, fill=arch_level_sub)) + geom_histogram(color="grey40") + ggtheme + xlim(-1,1) + xlab("Cor(mutation effect on temp,\nRDA-predicted temp. effect)") + guides(fill=guide_legend(title="Pleiotropy\nArchitecture")) 


ggplot(final.df, aes(x=RDA_RDAmutpred_cor_salEffect, fill=demog_level)) + geom_histogram(color="grey40") + ggtheme + xlim(0,1) + xlab("Cor(mutation effect on temp,\nRDA-predicted temp. effect)") + guides(fill=guide_legend(title="Landscape")) 

ggplot(final.df, aes(x=RDA_RDAmutpred_cor_salEffect, fill=demog_level_sub)) + geom_histogram(color="grey40") + ggtheme + xlim(-1,1) + xlab("Cor(mutation effect on temp,\nRDA-predicted temp. effect)") + guides(fill=guide_legend(title="Landscape"))

ggplot(final.df, aes(x=RDA_RDAmutpred_cor_salEffect, fill=arch_level_sub)) + geom_histogram(color="grey40") + ggtheme + xlim(-1,1) + xlab("Cor(mutation effect on temp,\nRDA-predicted temp. effect)") + guides(fill=guide_legend(title="Pleiotropy\nArchitecture")) 


```

Well, this is interesting. The accuracy of the mutation loadings on the RDA axis is much lower for the highly polygenic architectures. Which makes sense.

The ironic thing is that the phenotypic prediction from the RDA is still accurate, and it is affected more by demography. This suggests that the RDA prediction is reliable when the pattern of adaptation is correlated with genome-wide patterns of differentation, and not messed up by population structure.
```{r}
# STat models

summary(aov(cor_RDA20000_RDloadings_tempPhen~demog_level + demog_level_sub + arch_level, data=final.df))
summary(aov(cor_RDA20000_RDloadings_salPhen~demog_level + demog_level_sub + arch_level, data=final.df))

summary(aov(RDA_RDAmutpred_cor_tempEffect~demog_level + demog_level_sub + arch_level, data=final.df))
summary(aov(RDA_RDAmutpred_cor_salEffect~demog_level + demog_level_sub + arch_level, data=final.df))
```

Some additional RDA plots to show the effect of demography
```{r}

rda.df <- gather(final.df, key=trait, value=cor_RDApred_phen, cor_RDA20000_RDloadings_tempPhen, cor_RDA20000_RDloadings_salPhen)

r0 <- ggplot(rda.df, aes(x=cor_RDApred_phen, fill=demog_level)) + geom_histogram(color="grey30") + ggtheme + xlim(0,1) + xlab("Cor( phenotype, RDA predicted phenotype)") + facet_grid(~trait)

r1 <- ggplot(rda.df, aes(x=cor_RDApred_phen, fill=arch_level)) + geom_histogram(color="grey30") + ggtheme + xlim(0,1) + xlab("Cor( phenotype, RDA predicted phenotype)") + facet_grid(~trait)

r2 <- ggplot(rda.df, aes(x=cor_RDApred_phen, fill=as.factor(ispleiotropy))) + geom_histogram(color="grey30") + ggtheme + xlim(0,1) + xlab("Cor(temp. phenotype, RDA predicted temp. phenotype)") + facet_grid(~trait)

r3 <- ggplot(rda.df, aes(x=cor_RDApred_phen, fill=demog_level_sub)) + geom_histogram(color="grey30") + ggtheme + xlim(0,1) + xlab("Cor(temp. phenotype, RDA predicted temp. phenotype)") + facet_grid(~trait)

r4 <- ggplot(rda.df, aes(x=cor_RDApred_phen, fill=as.factor(SIGMA_K_2))) + geom_histogram(color="grey30") + ggtheme + xlim(0,1) + xlab("Cor(temp. phenotype, RDA predicted temp. phenotype)") + facet_grid(~trait)

pdf(paste0(outputs,"RDA_phen_predict_breakdown.pdf"), height=12, width=10)
grid.arrange(r0, r1, r2, r3, r4, nrow=5)
dev.off()

trait.labs <- c("Env2", "Temperature")
names(trait.labs) <- c("Env2", "Temperature")

r4 <- ggplot(final.df, aes(x=cor_RDA20000_RDloadings_tempPhen, fill=as.factor(SIGMA_K_2))) + geom_histogram(color="grey30") + ggtheme + xlim(0,1) + xlab("Cor(temp phenotype,\nRDA-predicted temp phenotype)") +  guides(fill=guide_legend(title="Stabilizing\nSelection"))

pdf(paste0(outputs,"RDA-predict-strengthselection.pdf"), width=4, height=3)
  r4
dev.off()
```

```{r}
#Scatterplot
  
n1 <- ggplot(final.df[order(final.df$demog_level_sub, decreasing=TRUE),]) +   geom_point(aes(x=subsamp_corr_phen_temp,
                   y=cor_RDA20000_RDloadings_tempPhen, color=demog_level,
	                   shape=demog_level_sub), bg="red",alpha=0.3) + ylim(0,1) + xlim(0,1) +
     ggtheme + geom_abline(slope=1, color="grey") + 
  scale_shape_manual(values = c(7, 10, 12, 21, 22), name="Demography") + guides(color=guide_legend(title="Landscape")) +
  xlab("Cor(Temp. phenotype, Deme Temp.)") +
  ylab("Cor(RDA prediction, Temp. phenotype)")
  
n1 

n2<-  ggplot(final.df[order(final.df$demog_level_sub, decreasing=TRUE),]) +   geom_point(aes(x=subsamp_corr_phen_sal,
                   y=cor_RDA20000_RDloadings_salPhen, color=demog_level,
	                   shape=demog_level_sub), alpha=0.3, bg="red") + ylim(0,1) + xlim(0,1) + 
     ggtheme + geom_abline(slope=1, color="grey") + 
  scale_shape_manual(values = c(7, 10, 12, 21, 22), name="Demography") + guides(color=guide_legend(title="Landscape")) +
  xlab("Cor(Env2 phenotype, Deme Env2)") +
  ylab("Cor(RDA prediction, Env2 phenotype)")
  

pdf(paste0(outputs,"RDA-Demog-scatter.pdf"), width=6, height=7)
  grid.arrange(n1, n2, nrow=2)
dev.off()


#hist(final.df$cor_RDA20000_RDloadings_tempPhen, xlim=c(0,1), breaks=seq(0,1,0.05))
#hist(final.df$cor_RDA20000_RDloadings_salPhen, xlim=c(0,1), breaks=seq(0,1,0.05))
```

